---
title: "Broadband Affordability Research"
author: "Bree Norlander"
date: "`r format(Sys.time(), '%B %d, %Y')`"
#output: html_notebook
---

Affordability research
We’d take data from the Commitments dataset and look at monthly cost to arrive at a per Mbps figures the primary affordability figure
We’d take this base measure across multiple program years for any library that received a commitment in the post-modernization period (2016 – 2023)
We could look at this through various filters / lenses:
Rural vs urban – in which locale is the average cost higher?
Discount bands – do higher discount bands pay more or less (pre-discount)?
Connection type – do fiber, copper, wireless, or other connection types have higher or lower average costs?
Application Type – do consortium vs library / library system applications have higher or lower costs?


```{r}
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(aws.s3)
library(janitor)
```

## Read in data:

```{r}
# Read in 2022 cat1 erate data
cat1_2022 <-
  s3read_using(
    FUN = read.csv,
    na.strings = c("", " ", "N/A", "n/a"),
    object = "2022_Libraries_Funded_Committed_Category_1.csv",
    bucket = "erate-data/data/AVI8-SVP9_Commitments"
  )

# Read in IMLS USAC matches
matches <-
  s3read_using(
    FUN = read.csv,
    na.strings = c("", " ", "N/A", "n/a"),
    object = "Algorithmic_Matches.csv",
    bucket = "erate-data/data/USAC_IMLS_Match"
  )

# Read in 2022 purpose erate data
purpose <-
  s3read_using(
    FUN = read.csv,
    na.strings = c("", " ", "N/A", "n/a"),
    object = "2022_Full_Purpose.csv",
    bucket = "erate-data/data/HBJ5-2BPJ_Purpose"
  )
```

```{r}
# Read in IMLS PLS AE 2021 dataset stored in S3
#imls_ae_2021 <- s3read_using(FUN = read.csv, object = "data/IMLS_PLS/2021_IMLS_PLS_AE.csv", bucket = "erate-data")

# Read in IMLS PLS AE 2021 dataset stored in S3
imls_outs <- s3read_using(FUN = read.csv, object = "data/IMLS_PLS/2014-2021_IMLS_Outlets_Unique.csv.csv", bucket = "erate-data")
```

```{r}
# Add in descriptions of locales
imls_outs <- imls_outs |> 
  mutate(
    LOCALE_DESCR = case_when(
      LOCALE == 11 ~ "City Large",
      LOCALE == 12 ~ "City Midsize",
      LOCALE == 13 ~ "City Small",
      LOCALE == 21 ~ "Suburban Large",
      LOCALE == 22 ~ "Suburban Midsize",
      LOCALE == 23 ~ "Suburban Small",
      LOCALE == 31 ~ "Town Fringe",
      LOCALE == 32 ~ "Town Distant",
      LOCALE == 33 ~ "Town Remote",
      LOCALE == 41 ~ "Rural Fringe",
      LOCALE == 42 ~ "Rural Distant",
      LOCALE == 43 ~ "Rural Remote"
    )) %>% 
  # separate to create high-level category
  # https://stackoverflow.com/a/53701998
  separate(LOCALE_DESCR, 
        into = c("LOCALE_TOP_LEVEL_DESCR", NA),
        sep = " ",
        remove = F)
```

# Add in the FSCSKEY and FSCS_SEQ to USAC data

```{r}
# Add in IMLS designators and normalize the upload/download speeds
cat1_2022_wrangled <- cat1_2022 |>
  left_join(matches |> select(FSCSKEY, FSCS_SEQ, ros_entity_number),
            by = join_by(ros_entity_number)) |>
  mutate(
    download_speed_mbps = case_when(
      form_471_download_speed_unit_name == "Mbps" ~ as.numeric(download_speed),
      form_471_download_speed_unit_name == "Gbps" ~ as.numeric(download_speed) * 1000
    )
  ) %>%
  mutate(
    upload_speed_mbps = case_when(
      form_471_upload_speed_unit_name == "Mbps" ~ as.numeric(upload_speed),
      form_471_upload_speed_unit_name == "Gbps" ~ as.numeric(upload_speed) * 1000
    )
  ) |>
  mutate(download_cost_per_mbps_per_month = case_when(
    !is.na(monthly_recurring_unit_eligible_costs) &&
      !is.na(download_speed_mbps) ~ as.numeric(monthly_recurring_unit_eligible_costs) /
      as.numeric(download_speed_mbps)
  ))
```

```{r}
# Create a dataframe of just the columns we want for analysis
afford_df <- cat1_2022_wrangled |> 
  filter(!is.na(FSCSKEY)) |> 
  distinct(ros_entity_number, FSCSKEY, FSCS_SEQ, ros_entity_name, ros_entity_type, ros_subtype, ros_physical_state, ros_urban_rural_status, billed_entity_number, organization_name, organization_entity_type_name, org_state, funding_year, application_number, form_471_line_item_number, form_471_function_name, form_471_product_name, total_monthly_cost, monthly_recurring_unit_eligible_costs, download_speed_mbps, upload_speed_mbps, download_cost_per_mbps_per_month) |> 
  filter(!is.na(download_speed_mbps),
         form_471_product_name != "Data plan for portable device",
         form_471_product_name != "Maintenance and Technical Support") |> 
  add_count(ros_entity_number, form_471_function_name, form_471_product_name, name="how_many_of_same_connection") |> 
  mutate(weight_for_mean = 1 / how_many_of_same_connection) # Create a weight for mean calc
  
```

```{r}
# Join cat1_2022 to purpose dataset
afford_df <- afford_df |> 
  mutate_at(c('application_number', 'form_471_line_item_number'), as.numeric) |> 
  left_join((purpose %>% 
               select(application_number, form_471_line_item_number, form_471_purpose_name, form_version, connection_directly_school, connection_supports_service) %>% 
               filter(form_version == "Current") %>% 
               mutate(application_number = as.numeric(application_number),
                      form_471_line_item_number = as.numeric(form_471_line_item_number))
             ), 
            by=c("application_number", "form_471_line_item_number")) 
```

```{r}
# write out csv for entities with 6 or more connections
# afford_df |> 
#   filter(how_many_of_same_connection > 5) |> 
#   write.csv("~/Downloads/More_than_5_connections.csv")
```


```{r}
mean(afford_df$download_cost_per_mbps_per_month, na.rm = TRUE)
```


```{r}
afford_df |> 
  #group_by(ros_urban_rural_status) |> 
  summarise(weighted.mean(x=download_cost_per_mbps_per_month, w=weight_for_mean, na.rm=TRUE))
```

```{r}
afford_df |> 
  group_by(ros_urban_rural_status) |> 
  summarise(weighted.mean(x=download_cost_per_mbps_per_month, w=weight_for_mean, na.rm=TRUE))
```

```{r}
afford_df |> 
  filter(download_speed_mbps < 500000) |> 
  ggplot(aes(x=download_cost_per_mbps_per_month, y=download_speed_mbps)) +
  geom_point()
```

```{r}
afford_df |> 
  filter(download_speed_mbps > 500000)
```

```{r}
afford_df |> 
  filter(download_speed_mbps < 500000) |> 
  ggplot(aes(x=form_471_function_name, y=download_cost_per_mbps_per_month)) +
  geom_point()
```

```{r}
afford_df |> 
  filter(download_speed_mbps < 500000) |> 
  ggplot(aes(x=form_471_function_name, y=download_speed_mbps)) +
  geom_point()
```

```{r}
afford_df |> 
  filter(download_speed_mbps < 500000,
         form_471_purpose_name != "Backbone circuit for consortium that provides connectivity between aggregation points or other non-user facilities") |> 
  ggplot(aes(x=form_471_function_name, y=download_speed_mbps)) +
  geom_point()
Backbone circuit for consortium that provides connectivity between aggregation points or other non-user facilities + Applicant owned fiber network, eligible fiber electronics, or maintenance & operations
```





