---
title: "E-Rate Analysis"
output: html_notebook
---

```{r}
# turn off scientific notation
options(scipen=999)
# show specific number of digits
options(digits=3)
```

```{r}
# Load libraries
library(RSocrata)
library(rquery)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(glue)
library(purrr)
library(janitor)
library(aws.s3)
library(gt)
#library(tidycensus)
```

```{r}
# Find out what distinct years exist in the dataset
read.csv("https://opendata.usac.org/resource/avi8-svp9.csv?$select=distinct%20funding_year")
```

```{r}
# Load datasets
disbursement <- s3read_using(FUN = read.csv, object = "data/QDMP-YGFT_Disbursements/Full_Funded_Disbursements.csv", bucket = "erate-data")

disbursement_merged <- s3read_using(FUN = read.csv, object = "data/USAC_Tables_Merged/Commitments_AVI8-SVP9_and_Disbursements_QDMP-YGFT.csv", bucket = "erate-data")

erate_libs <- s3read_using(FUN = read.csv, object = "data/AVI8-SVP9_Commitments/Libraries_Funded_Committed_AVI8-SVP9.csv", bucket = "erate-data")

erate_imls <- s3read_using(FUN = read.csv, object = "data/AVI8-SVP9_Commitments/Libraries_Funded_Committed_AVI8-SVP9_with_FSCS_Matches.csv", bucket = "erate-data")

ae_and_outlets <- s3read_using(FUN = read.csv, object = "data/IMLS_PLS/2014-2020_OUTLETs_and_AEs_Merged.csv", bucket = "erate-data")

# imls_outlet_erate <- s3read_using(FUN = read.csv, object = "data/imls_outlet_erate.csv", bucket = "erate-data")
# imls_outlet_erate <- imls_outlet_erate %>% select(-X)

# imls_ae_erate <- s3read_using(FUN = read.csv, object = "data/imls_ae_erate.csv", bucket = "erate-data")
# imls_ae_erate <- imls_ae_erate %>% select(-X)

state_pops <- s3read_using(FUN = read.csv, object = "data/Population_Data/State_Populations_2016-2021.csv", bucket = "erate-data")

# full_imls_2018 <- read.csv("~/Google Drive/Shared drives/ALA-ERate/Data/2021-06-03_IMLS_MashUp_2018.csv")

# full_imls_2019 <- read.csv("~/Google Drive/Shared drives/ALA-ERate/Data/2021-06-03_IMLS_MashUp_2019.csv")
```

```{r}
# Look at branches that are part of the consortium with ben 17023688
erate_imls %>% 
  filter(billed_entity_number == 17023688) %>% 
  select(ros_entity_number,
         ros_entity_name,
         ros_physical_state,
         organization_name,
         funding_year,
         chosen_category_of_service,
         cat2_discount_by_ros,
         cat1_discount_by_ros_estimated) %>%
  group_by(funding_year, ros_entity_number, ros_entity_name, ros_physical_state) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(Category1 = if(all(is.na(cat1_discount_by_ros_estimated))) NA_real_ 
            else round(sum(cat1_discount_by_ros_estimated, na.rm = T), 2),
            Category2 = if(all(is.na(cat2_discount_by_ros))) NA_real_ 
            else round(sum(cat2_discount_by_ros, na.rm = T), 2)) %>% 
  pivot_wider(names_from = c(funding_year), values_from = c(Category1, Category2))
```

To answer Chris J's question:
Download Speed = greater than or equal to 1000; Download Speed Unit = Mbps
AND Download Speed = greater than or equal to 1; Download Speed Unit = Gbps

```{r}
# How many entities requested download speeds of 1 Gbps or greater in 2019?
erate_libs %>%
  mutate(download_speed = as.numeric(download_speed)) %>%
  filter(funding_year == "2019") %>%
  filter(form_471_download_speed_unit_name == "Gbps" & download_speed >= 1 |
           form_471_download_speed_unit_name == "Mbps" & download_speed >= 1000) %>%
   summarise(n_distinct(ros_entity_number, na.rm = T))
```

```{r}
# How many entities requested download/upload speeds of 1 Gbps or greater per year?
erate_libs %>%
  mutate(download_speed = as.numeric(download_speed)) %>%
  group_by(funding_year) %>%
  filter(form_471_download_speed_unit_name == "Gbps" & download_speed >= 1 |
           form_471_download_speed_unit_name == "Mbps" & download_speed >= 1000) %>%
  summarise(Cnt_Entities_Gbps_Download = n_distinct(ros_entity_number, na.rm = T)) %>% 
  # Add in the upload counts
  left_join(erate_libs %>% 
              mutate(upload_speed = as.numeric(upload_speed)) %>%
              group_by(funding_year) %>%
              filter(form_471_upload_speed_unit_name == "Gbps" & upload_speed >= 1 | 
                       form_471_upload_speed_unit_name == "Mbps" & upload_speed >= 1000) %>%
              summarise(Cnt_Entities_Gbps_Upload = n_distinct(ros_entity_number, na.rm = T))) 
```

```{r}
# How many entities requested download speeds of 1 Gbps or greater per year?
erate_libs %>%
  mutate(download_speed = as.numeric(download_speed)) %>%
  group_by(funding_year) %>%
  filter(form_471_download_speed_unit_name == "Gbps" & download_speed >= 1 |
           form_471_download_speed_unit_name == "Mbps" & download_speed >= 1000) %>%
  summarise(Cnt_Entities_Gbps_Download = n_distinct(ros_entity_number, na.rm = T)) %>% 
  # Add in the upload counts
  left_join(erate_libs %>% 
              mutate(upload_speed = as.numeric(upload_speed)) %>%
              group_by(funding_year) %>%
              filter(form_471_upload_speed_unit_name == "Gbps" & upload_speed >= 1 | 
                       form_471_upload_speed_unit_name == "Mbps" & upload_speed >= 1000) %>%
              summarise(Cnt_Entities_Gbps_Upload = n_distinct(ros_entity_number, na.rm = T))) %>% 
  pivot_longer(cols = c(Cnt_Entities_Gbps_Download, Cnt_Entities_Gbps_Upload), names_to = "Type", values_to = "Count") %>%
  ggplot(aes(x=funding_year, y = Count, color = Type)) +
  geom_point() +
  geom_line()+
  theme_linedraw() +
  scale_color_manual(values = c("#3778bf", "#b0054b"), labels = c("Download Speed", "Upload Speed")) +
  labs(title = "Count of Library Entities Requesting\n1 Gbps or Greater Upload/Download Speeds",
       x = "Funding Year",
       y = "Count of Library Entities",
       color = "Speed Type")

#ggsave("~/Google Drive/Shared drives/ALA-ERate/Plots/Entities_Requesting_1Gpbs_or_Greater.png")
```

```{r}
# How many imls_outlets requested this by year?
imls_outlet_erate %>%
  mutate(download_speed = as.numeric(download_speed)) %>%
  group_by(funding_year) %>%
  filter(form_471_download_speed_unit_name == "Gbps" & download_speed >= 1 |
           form_471_download_speed_unit_name == "Mbps" & download_speed >= 1000) %>%
  summarise(Cnt_Entities_Gbps_Download = n_distinct(ros_entity_number, na.rm = T)) %>% 
  # Add in the upload counts
  left_join(imls_outlet_erate %>% 
              mutate(upload_speed = as.numeric(upload_speed)) %>%
              group_by(funding_year) %>%
              filter(form_471_upload_speed_unit_name == "Gbps" & upload_speed >= 1 | 
                       form_471_upload_speed_unit_name == "Mbps" & upload_speed >= 1000) %>%
              summarise(Cnt_Entities_Gbps_Upload = n_distinct(ros_entity_number, na.rm = T))) 
  
```

```{r}
# How many entities per state requested this in 2019?
erate_libs %>%
  mutate(download_speed = as.numeric(download_speed)) %>%
  filter(funding_year == "2019") %>%
  filter(form_471_download_speed_unit_name == "Gbps" & download_speed >= 1 |
           form_471_download_speed_unit_name == "Mbps" & download_speed >= 1000) %>%
  group_by(ros_physical_state) %>%
  summarise(Cnt_Entities_Gbps_Download = n_distinct(ros_entity_number, na.rm = T)) %>% 
  # Add in the upload counts
  left_join(erate_libs %>%
              mutate(upload_speed = as.numeric(upload_speed)) %>%
              filter(funding_year == "2019") %>%
              filter(form_471_upload_speed_unit_name == "Gbps" & upload_speed >= 1 |
                       form_471_upload_speed_unit_name == "Mbps" & upload_speed >= 1000) %>%
              group_by(ros_physical_state) %>%
              summarise(Cnt_Entities_Gbps_Upload = n_distinct(ros_entity_number, na.rm = T)))
```

could you do this in the analysis - there are fields that i think are the same from one fy to the next around bandwidth requested - could you track changes in an entity - so in 2016 this entity requested 100MB of bandwidth, in 2017 they requested the same, and then in 2018 they requested 1GB of bandwidth

```{r}
# Download speed requested over time by entity
erate_libs %>%
  mutate(download_speed = as.numeric(download_speed),
         ros_physical_state = str_to_lower(str_trim(ros_physical_state, side = "both")),
         ros_physical_city = str_to_lower(str_trim(ros_physical_city, side = "both"))) %>%
  # Normalize the download and upload speeds to Megabits per second
  mutate(download_speed_mbps = case_when(
    form_471_download_speed_unit_name == "Mbps" ~ download_speed,
    form_471_download_speed_unit_name == "Gbps" ~ download_speed * 1000
  )) %>%
  mutate(upload_speed_mbps = case_when(
    form_471_upload_speed_unit_name == "Mbps" ~ download_speed,
    form_471_upload_speed_unit_name == "Gbps" ~ download_speed * 1000
  )) %>% 
  group_by(ros_entity_number, ros_entity_name, funding_year) %>%
  arrange(funding_year) %>% 
  slice_max(download_speed_mbps, with_ties = F) %>% 
  select(ros_entity_number, ros_entity_name, ros_physical_city, ros_physical_state, funding_year, download_speed_mbps) %>% 
  pivot_wider(names_from = funding_year, values_from = download_speed_mbps) #%>%
  #write.csv("~/Google Drive/Shared drives/ALA-ERate/Data/Download_Mbps_Requested_over_Time_by_Entity.csv")
```

```{r}
# Upload speed requested over time by entity
erate_libs %>%
  mutate(download_speed = as.numeric(download_speed),
         ros_physical_state = str_to_lower(str_trim(ros_physical_state, side = "both")),
         ros_physical_city = str_to_lower(str_trim(ros_physical_city, side = "both"))) %>%
  # Normalize the download and upload speeds to Megabits per second
  mutate(download_speed_mbps = case_when(
    form_471_download_speed_unit_name == "Mbps" ~ download_speed,
    form_471_download_speed_unit_name == "Gbps" ~ download_speed * 1000
  )) %>%
  mutate(upload_speed_mbps = case_when(
    form_471_upload_speed_unit_name == "Mbps" ~ download_speed,
    form_471_upload_speed_unit_name == "Gbps" ~ download_speed * 1000
  )) %>% 
  group_by(ros_entity_number, ros_entity_name, funding_year) %>%
  arrange(funding_year) %>% 
  slice_max(upload_speed_mbps, with_ties = F) %>% 
  select(ros_entity_number, ros_entity_name, ros_physical_city, ros_physical_state, funding_year, upload_speed_mbps) %>% 
  pivot_wider(names_from = funding_year, values_from = upload_speed_mbps) #%>%
  #write.csv("~/Google Drive/Shared drives/ALA-ERate/Data/Upload_Mbps_Requested_over_Time_by_Entity.csv")
```

```{r}
# What is the average price per Mb download speed
# Upload speed requested over time by entity
erate_libs %>%
  mutate(download_speed = as.numeric(download_speed),
         upload_speed = as.numeric(upload_speed),
         ros_physical_state = str_to_lower(str_trim(ros_physical_state, side = "both")),
         ros_physical_city = str_to_lower(str_trim(ros_physical_city, side = "both"))) %>%
  # Normalize the download and upload speeds to Megabits per second
  mutate(download_speed_mbps = case_when(
    form_471_download_speed_unit_name == "Mbps" ~ download_speed,
    form_471_download_speed_unit_name == "Gbps" ~ download_speed * 1000
  )) %>%
  mutate(upload_speed_mbps = case_when(
    form_471_upload_speed_unit_name == "Mbps" ~ upload_speed,
    form_471_upload_speed_unit_name == "Gbps" ~ upload_speed * 1000
  )) %>% 
  mutate(download_cost_per_Mb = download_speed_mbps/total_monthly_cost,
         upload_cost_per_Mb = upload_speed_mbps/total_monthly_cost) %>% 
  group_by(funding_year) %>% 
  summarise(average_download_cost = mean(download_cost_per_Mb, trim = 0.05, na.rm = T),
    average_upload_cost = mean(upload_cost_per_Mb, trim = 0.05, na.rm = T))
```

# Playing around with possible dashboard visualizations

```{r}
# For iconference poster number of unique entities per year
erate_libs %>% 
  group_by(funding_year) %>% 
  summarise(n_distinct(ros_entity_number))
```

```{r}
# For number of unique entities per state per year
erate_imls %>% 
  group_by(funding_year, ros_physical_state) %>% 
  summarise(n_distinct(ros_entity_number))
```


```{r}
# For iconference poster number of unique imls entities per year
erate_imls %>% 
  filter(!is.na(FSCSKEY) & !is.na(FSCS_SEQ)) %>% 
  group_by(funding_year) %>% 
  summarise(n_distinct(ros_entity_number))
```

```{r}
# Create a table of cat1 and cat 2 discounts (that could be filtered by year)
# https://stackoverflow.com/a/54805598/5593458 This is about NA summing
# also https://stackoverflow.com/a/24232804/5593458
erate_libs %>% 
  select(ros_physical_state, funding_year, chosen_category_of_service, cat2_discount_by_ros, cat1_discount_by_ros_estimated) %>%
  filter(funding_year == "2019") %>%
  group_by(ros_physical_state) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(sum_cat1_discount = if(all(is.na(cat1_discount_by_ros_estimated))) NA_real_ 
            else sum(cat1_discount_by_ros_estimated, na.rm = T),
            sum_cat2_discount = if(all(is.na(cat2_discount_by_ros))) NA_real_ 
            else sum(cat2_discount_by_ros, na.rm = T)) %>%
  # Add a columns showing state total AND the Pct of libraries total
  mutate(state_total = rowSums(select(., sum_cat1_discount, sum_cat2_discount), na.rm = TRUE),
         pct_of_library_total = paste0(round(state_total / sum(state_total, na.rm = T)*100, 2), "%")) 
```

```{r}
# Create a table of cat1 and cat 2 discounts (that could be filtered by year)
# https://stackoverflow.com/a/54805598/5593458 This is about NA summing
# also https://stackoverflow.com/a/24232804/5593458
erate_imls %>% 
  select(ros_physical_state, funding_year, chosen_category_of_service, cat2_discount_by_ros, cat1_discount_by_ros_estimated) %>%
  filter(ros_physical_state == 'WA') %>% 
  group_by(funding_year) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(sum_cat1_discount = if(all(is.na(cat1_discount_by_ros_estimated))) NA_real_ 
            else sum(cat1_discount_by_ros_estimated, na.rm = T),
            sum_cat2_discount = if(all(is.na(cat2_discount_by_ros))) NA_real_ 
            else sum(cat2_discount_by_ros, na.rm = T)) %>%
  rowwise() %>% 
  # Add a columns showing state total 
  mutate(state_total = sum(c(sum_cat1_discount, sum_cat2_discount), na.rm = TRUE)) 
```



```{r}
# Create a table of cat1 and cat 2 discounts using only imls_outlet data
# https://stackoverflow.com/a/54805598/5593458 This is about NA summing
# also https://stackoverflow.com/a/24232804/5593458
imls_outlet_erate %>% # just IMLS outlets
  select(ros_physical_state, funding_year, chosen_category_of_service, cat2_discount_by_ros, cat1_discount_by_ros_estimated) %>%
  filter(funding_year == "2019") %>%
  group_by(ros_physical_state) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(sum_cat1_discount = if(all(is.na(cat1_discount_by_ros_estimated))) NA_real_ 
            else sum(cat1_discount_by_ros_estimated, na.rm = T),
            sum_cat2_discount = if(all(is.na(cat2_discount_by_ros))) NA_real_ 
            else sum(cat2_discount_by_ros, na.rm = T)) %>%
  # Add a columns showing state total AND the Pct of libraries total
  mutate(state_total = rowSums(select(., sum_cat1_discount, sum_cat2_discount), na.rm = TRUE),
         pct_of_library_total = paste0(round(state_total / sum(state_total, na.rm = T)*100, 2), "%")) 
```

```{r}
# library(tidycensus)
# # Create a dataframe of state populations
# years <- read.csv("https://opendata.usac.org/resource/avi8-svp9.csv?$select=distinct%20funding_year")[[1]]
# 
# y = list()
# 
# # Call the API for each year's population statistics (but there is no data for 2020 in this location (or 2021) hence minus 2)
# for (i in 1:(length(years)-2)) {
#   y[[i]] <- get_estimates(geography = "state",
#                      product = "population",
#                      year = years[i],
#                      key = Sys.getenv("CENSUS_API_KEY"))
# 
#   y[[i]] <- y[[i]] %>%
#     mutate(year = years[i]) %>%
#     filter(variable == "POP") %>%
#     mutate(STABR = case_when(
#            NAME != "Puerto Rico" ~ usdata::state2abbr(NAME),
#            NAME == "Puerto Rico" ~ "PR"
#          ))
# }
# 
# state_pops <- bind_rows(y)
# 
# # 2020 pops retrieved from https://www.census.gov/data/tables/2020/dec/2020-apportionment-data.html
# pop_2020 <- readxl::read_xlsx("~/Google Drive/Shared drives/ALA-ERate/Data/CENSUS-apportionment-2020-table02.xlsx", range = "A4:B57") %>% 
#   slice(-52L) %>% 
#   mutate(STABR = case_when(
#            AREA != "Puerto Rico" ~ usdata::state2abbr(AREA),
#            AREA == "Puerto Rico" ~ "PR"
#          )) %>% 
#   mutate(variable = "POP",
#          year = 2020) %>% 
#   rename(NAME = AREA,
#          value = `RESIDENT POPULATION (APRIL 1, 2020)`) %>% 
#   left_join(state_pops %>% distinct(STABR, GEOID))
# 
# state_pops <- state_pops %>% rbind(pop_2020) 
# 
# write.csv(state_pops, "~/Google Drive/Shared drives/ALA-ERate/Data/State_Populations_2016-2020.csv")
```

```{r}
# Per capita calculations
# Create a table of cat1 and cat 2 discounts
# https://stackoverflow.com/a/54805598/5593458 This is about NA summing
# also https://stackoverflow.com/a/24232804/5593458
erate_libs %>% 
  select(ros_physical_state, funding_year, chosen_category_of_service, cat2_discount_by_ros, cat1_discount_by_ros_estimated) %>%
  filter(funding_year == "2019") %>%
  group_by(ros_physical_state) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(sum_cat1_discount = if(all(is.na(cat1_discount_by_ros_estimated))) NA_real_ 
            else sum(cat1_discount_by_ros_estimated, na.rm = T),
            sum_cat2_discount = if(all(is.na(cat2_discount_by_ros))) NA_real_ 
            else sum(cat2_discount_by_ros, na.rm = T)) %>%
  # Add a columns showing state total AND the Pct of libraries total
  mutate(state_total = rowSums(select(., sum_cat1_discount, sum_cat2_discount), na.rm = TRUE),
         pct_of_library_total = paste0(round(state_total / sum(state_total, na.rm = T)*100, 2), "%")) %>% 
  as.data.frame() %>% 
  # add in populations stats and calculate per capita discount
  left_join(state_pops %>% 
              filter(year == "2019") %>% 
              select(value, STABR), 
            by = c("ros_physical_state" = "STABR")
            ) %>% 
  rename(state_pop = value) %>% 
  mutate(per_capita_discount = round(state_total / state_pop, 2)) %>% 
  arrange(desc(per_capita_discount))
```

```{r}
# Per capita calculations
# Create a table of cat1 and cat 2 discounts using only imls outlets data
# https://stackoverflow.com/a/54805598/5593458 This is about NA summing
# also https://stackoverflow.com/a/24232804/5593458
imls_outlet_erate %>%  # IMLS outlets only
  select(ros_physical_state, funding_year, chosen_category_of_service, cat2_discount_by_ros, cat1_discount_by_ros_estimated) %>%
  filter(funding_year == "2019") %>%
  group_by(ros_physical_state) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(sum_cat1_discount = if(all(is.na(cat1_discount_by_ros_estimated))) NA_real_ 
            else sum(cat1_discount_by_ros_estimated, na.rm = T),
            sum_cat2_discount = if(all(is.na(cat2_discount_by_ros))) NA_real_ 
            else sum(cat2_discount_by_ros, na.rm = T)) %>%
  # Add a columns showing state total AND the Pct of libraries total
  mutate(state_total = rowSums(select(., sum_cat1_discount, sum_cat2_discount), na.rm = TRUE),
         pct_of_library_total = paste0(round(state_total / sum(state_total, na.rm = T)*100, 2), "%")) %>% 
  as.data.frame() %>% 
  # add in populations stats and calculate per capita discount
  left_join(state_pops %>% 
              filter(year == "2019") %>% 
              select(value, STABR), 
            by = c("ros_physical_state" = "STABR")
            ) %>% 
  rename(state_pop = value) %>% 
  mutate(per_capita_discount = round(state_total / state_pop, 2)) %>% 
  arrange(desc(per_capita_discount))
```

```{r}
# Specific csv request from CJ


# Per capita calculations
# Create a table of cat1 and cat 2 discounts using only imls outlets data
# https://stackoverflow.com/a/54805598/5593458 This is about NA summing
# also https://stackoverflow.com/a/24232804/5593458
imls_outlet_erate %>%  # IMLS outlets only
  select(ros_physical_state, funding_year, chosen_category_of_service, cat2_discount_by_ros, cat1_discount_by_ros_estimated) %>%
  group_by(funding_year, ros_physical_state) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(sum_cat1_discount = if(all(is.na(cat1_discount_by_ros_estimated))) NA_real_ 
            else sum(cat1_discount_by_ros_estimated, na.rm = T),
            sum_cat2_discount = if(all(is.na(cat2_discount_by_ros))) NA_real_ 
            else sum(cat2_discount_by_ros, na.rm = T)) %>%
  # Add a columns showing state total AND the Pct of libraries total
  rowwise() %>% 
  mutate(state_year_total = sum(sum_cat1_discount, sum_cat2_discount, na.rm = TRUE)) %>% 
  as.data.frame() %>% 
  # add in populations stats and calculate per capita discount
  left_join(state_pops %>% 
              select(value, STABR, year), 
            by = c("ros_physical_state" = "STABR", "funding_year" = "year")
            ) %>% 
  rename(state_pop = value) %>% 
  mutate(cat1_per_capita_discount = round(sum_cat1_discount / state_pop, 2),
         cat2_per_capita_discount = round(sum_cat2_discount / state_pop, 2),
         total_per_capita_discount = round(state_year_total / state_pop, 2)) #%>% 
  #write.csv((paste0("~/Google Drive File Stream/Shared drives/ALA-ERate/Data/States_Years_Cat1_Cat2_", Sys.Date(), ".csv")))
```

```{r}
# Use only IMLS recipients
# Create a table of cat1 and cat 2 discounts & per_capita_discounts by year and state
# https://stackoverflow.com/a/54805598/5593458 This is about NA summing
# also https://stackoverflow.com/a/24232804/5593458
erate_imls %>% 
  filter(!is.na(FSCSKEY)) %>% # Just IMLS entities
  select(ros_physical_state, funding_year, chosen_category_of_service, cat2_discount_by_ros, cat1_discount_by_ros_estimated) %>%
  group_by(ros_physical_state, funding_year) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(sum_cat1_discount = if(all(is.na(cat1_discount_by_ros_estimated))) NA_real_ 
            else sum(cat1_discount_by_ros_estimated, na.rm = T),
            sum_cat2_discount = if(all(is.na(cat2_discount_by_ros))) NA_real_ 
            else sum(cat2_discount_by_ros, na.rm = T)) %>%
  # Add a columns showing state total AND the Pct of libraries total
  rowwise() %>% 
  mutate(state_total = sum(c(sum_cat1_discount, sum_cat2_discount), na.rm = TRUE)) %>% 
  as.data.frame() %>% 
  # add in populations stats and calculate per capita discount
  left_join(state_pops %>% 
              select(value, year, STABR), 
            by = c("ros_physical_state" = "STABR", "funding_year" = "year")
            ) %>% 
  rename(state_pop = value) %>% 
  mutate(per_capita_discount = round(state_total / state_pop, 2))
```

```{r}
# Create a table of cat1 and cat 2 discounts & per_capita_discounts by year and state
# https://stackoverflow.com/a/54805598/5593458 This is about NA summing
# also https://stackoverflow.com/a/24232804/5593458
erate_imls %>% 
  filter(!is.na(FSCSKEY)) %>% # Just IMLS entities
  select(ros_physical_state, funding_year, chosen_category_of_service, cat2_discount_by_ros, cat1_discount_by_ros_estimated) %>%
  group_by(ros_physical_state, funding_year) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(sum_cat1_discount = if(all(is.na(cat1_discount_by_ros_estimated))) NA_real_ 
            else sum(cat1_discount_by_ros_estimated, na.rm = T),
            sum_cat2_discount = if(all(is.na(cat2_discount_by_ros))) NA_real_ 
            else sum(cat2_discount_by_ros, na.rm = T)) %>%
  # Add a columns showing state total AND the Pct of libraries total
  rowwise() %>% 
  mutate(state_total = sum(c(sum_cat1_discount, sum_cat2_discount), na.rm = TRUE)) %>% 
  as.data.frame() %>% 
  # add in populations stats and calculate per capita discount
  left_join(state_pops %>% 
              select(value, year, STABR), 
            by = c("ros_physical_state" = "STABR", "funding_year" = "year")
            ) %>% 
  rename(state_pop = value) %>% 
  mutate(per_capita_discount = round(state_total / state_pop, 2)) %>% 
  select(-sum_cat1_discount, -sum_cat2_discount, -state_total, -state_pop) %>% 
  pivot_wider(names_from = funding_year, values_from = per_capita_discount)
```

```{r}
# Create a table of cat1 and cat 2 discounts & per_capita_discounts by year and state
# https://stackoverflow.com/a/54805598/5593458 This is about NA summing
# also https://stackoverflow.com/a/24232804/5593458
erate_imls %>%
  filter(!is.na(FSCSKEY)) %>% # Just IMLS entities
  select(
    ros_physical_state,
    funding_year,
    chosen_category_of_service,
    cat2_discount_by_ros,
    cat1_discount_by_ros_estimated
  ) %>%
  group_by(ros_physical_state, funding_year) %>%
  summarise(
    sum_cat1_discount = if (all(is.na(cat1_discount_by_ros_estimated)))
      NA_real_
    else
      sum(cat1_discount_by_ros_estimated, na.rm = T),
    sum_cat2_discount = if (all(is.na(cat2_discount_by_ros)))
      NA_real_
    else
      sum(cat2_discount_by_ros, na.rm = T)
  ) %>%
  rowwise() %>%
  mutate(state_total = sum(c(sum_cat1_discount, sum_cat2_discount), na.rm = TRUE)) %>%
  as.data.frame() %>%
  # add in populations stats and calculate per capita discount
  left_join(
    state_pops %>% select(value, year, STABR),
    by = c("ros_physical_state" = "STABR",
           "funding_year" = "year")
  ) %>%
  rename(state_pop = value) %>%
  mutate(per_capita_discount = round(state_total / state_pop, 2)) %>%
  filter(ros_physical_state == "AK") %>%
  ggplot(aes(x = funding_year, y = per_capita_discount)) +
  geom_col(
    color = "black",
    fill = "#3778bf",
    size = 0.2,
    position = "dodge"
  ) +
  theme_linedraw() +
  geom_text(
    x = 2020,
    y = 0,
    hjust = -0.3,
    label = "No Data Available",
    color = "black",
    size = 3.5,
    angle = 90
  ) +
  scale_y_continuous(labels = scales::dollar_format()) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  labs(title = paste0("Per Capita Commitment"),
       x = "Funding Year",
       y = "Commitment Per Capita")

```

```{r}
# Create a table of per entity discounts by year and state
# https://stackoverflow.com/a/54805598/5593458 This is about NA summing
# also https://stackoverflow.com/a/24232804/5593458
erate_imls %>%
  filter(!is.na(FSCSKEY)) %>% # Just IMLS entities
  select(ros_entity_number,
    ros_physical_state,
    funding_year,
    chosen_category_of_service,
    cat2_discount_by_ros,
    cat1_discount_by_ros_estimated
  ) %>%
  group_by(ros_physical_state, funding_year) %>%
  summarise(
    sum_cat1_discount = if (all(is.na(cat1_discount_by_ros_estimated)))
      NA_real_
    else
      sum(cat1_discount_by_ros_estimated, na.rm = T),
    sum_cat2_discount = if (all(is.na(cat2_discount_by_ros)))
      NA_real_
    else
      sum(cat2_discount_by_ros, na.rm = T),
    cnt_entities = n_distinct(ros_entity_number)
  ) %>%
  rowwise() %>%
  mutate(state_total = sum(c(sum_cat1_discount, sum_cat2_discount), na.rm = TRUE),
         per_entity = state_total/cnt_entities) 
```

```{r}
# Create a table of commitments per_square footage by year and state
# https://stackoverflow.com/a/54805598/5593458 This is about NA summing
# also https://stackoverflow.com/a/24232804/5593458
erate_imls %>% 
  filter(!is.na(FSCSKEY) & !is.na(FSCS_SEQ)) %>%  # Just IMLS entities
  select(ros_entity_number, ros_physical_state, funding_year, chosen_category_of_service, cat2_discount_by_ros, cat1_discount_by_ros_estimated, ros_square_footage) %>%
  group_by(ros_physical_state, funding_year) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(sum_cat1_discount = if(all(is.na(cat1_discount_by_ros_estimated))) NA_real_ 
            else sum(cat1_discount_by_ros_estimated, na.rm = T),
            sum_cat2_discount = if(all(is.na(cat2_discount_by_ros))) NA_real_ 
            else sum(cat2_discount_by_ros, na.rm = T)) %>%
  # Add a column showing state total
  rowwise() %>% 
  mutate(state_total = sum(c(sum_cat1_discount, sum_cat2_discount), na.rm = TRUE)) %>% 
  left_join(erate_imls %>%
              filter(!is.na(FSCSKEY) & !is.na(FSCS_SEQ),
                     ros_square_footage > 0 & SQ_FEET > 0) %>%  
              group_by(ros_entity_number, funding_year) %>%
              slice(n=1) %>%
              ungroup() %>% 
              group_by(ros_physical_state, funding_year) %>% 
              summarise(sum_sq_ft = sum(ros_square_footage, na.rm = T))) %>% 
  as.data.frame() %>% 
  mutate(per_sq_ft_commit = round(state_total / sum_sq_ft, 2)) %>% 
  select(-sum_cat1_discount, -sum_cat2_discount, -state_total, -sum_sq_ft) %>% 
  pivot_wider(names_from = funding_year, values_from = per_sq_ft_commit)
```

```{r}
erate_imls %>%
  filter(!is.na(ros_entity_number) & !is.na(FSCS_SEQ) & !is.na(STABR),
         ros_square_footage > 0 & SQ_FEET > 0) %>% 
  distinct(ros_entity_number, funding_year, .keep_all = T) %>%
  group_by(ros_entity_number) %>% 
  slice_max(funding_year, n=1) %>%
  ungroup() %>% 
  group_by(STABR, .drop = F) %>% 
  summarise(Erate_Square_Footage = sum(ros_square_footage, na.rm = T),
            IMLS_Square_Footage = sum(SQ_FEET, na.rm = T)) #%>% 
  #write.csv((paste0("~/Google Drive/Shared drives/ALA-ERate/Data/Square_Footage_Comparison_", Sys.Date(), ".csv")))
```

```{r}
# Create a table with national Cat1 and Cat2 totals by year
erate_libs %>%
  select(funding_year,
         chosen_category_of_service,
         cat2_discount_by_ros,
         cat1_discount_by_ros_estimated) %>%
  group_by(funding_year) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(Category1_Commitment = if(all(is.na(cat1_discount_by_ros_estimated))) NA_real_ 
            else sum(cat1_discount_by_ros_estimated, na.rm = T),
            Category2_Commitment = if(all(is.na(cat2_discount_by_ros))) NA_real_ 
            else sum(cat2_discount_by_ros, na.rm = T)) %>%
  rowwise() %>% 
  mutate(National_Total = sum(c(Category1_Commitment, Category2_Commitment), na.rm = TRUE)) %>% 
  gt() %>% #https://themockup.blog/posts/2020-09-04-10-table-rules-in-r/
  fmt_currency(
    columns = vars(`Category1_Commitment`, `Category2_Commitment`, `National_Total`), # reference cols by name
    currency = "USD"
  ) %>% 
  cols_align(align = "left",
             columns = 'funding_year') %>% 
  cols_align(align = "right",
             columns = vars(`Category1_Commitment`, `Category2_Commitment`, `National_Total`)) %>% 
  cols_width(vars(funding_year) ~ px(50),
             vars(Category1_Commitment, Category2_Commitment, National_Total) ~ px(170)
  ) %>% 
  cols_label(
    funding_year = md("Year"),
    'Category1_Commitment' = md("Category 1<br>Commitment"),
    'Category2_Commitment' = md("Category 2<br>Commitment"),
    'National_Total' = md("Total<br>Commitment")
  ) %>% 
  tab_header(
    title = md("**National Summary of Library Entities**")
    ) %>% 
  tab_style(
    style = cell_fill(
      color = "#DCDCDC"
    ),
    locations = list(
      cells_body(
        columns = vars(National_Total)
        )
      )
      ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_column_labels(everything())
  ) %>% 
  tab_style(
    style = list(
      cell_borders(
        sides = "left",
        color = "black",
        weight = px(3)
      )
    ),
    locations = list(
      cells_body(
        columns = vars(National_Total)
      )
    )
    ) %>% 
  opt_table_lines(extent = "default") %>%
  tab_options(
    heading.align = "left",
    table.border.top.color = "#3778bf",
    table.border.top.width = px(3),
    column_labels.border.top.color = "#3778bf",
    column_labels.border.top.width = px(3),
    column_labels.border.bottom.color = "black",
    table_body.hlines.color = "black",
    table.border.bottom.color = "#3778bf",
    table.border.bottom.width = px(3),
    data_row.padding = px(12)
  ) 

```

```{r}
# Create a table with national Cat1 and Cat2 totals & percents by year
erate_libs %>%
  select(funding_year,
         chosen_category_of_service,
         cat2_discount_by_ros,
         cat1_discount_by_ros_estimated) %>%
  group_by(funding_year) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(Category1 = if(all(is.na(cat1_discount_by_ros_estimated))) NA_real_ 
            else sum(cat1_discount_by_ros_estimated, na.rm = T),
            Category2 = if(all(is.na(cat2_discount_by_ros))) NA_real_ 
            else sum(cat2_discount_by_ros, na.rm = T)) %>%
  pivot_longer(!funding_year, names_to = "Category", values_to = "tot") %>% 
  pivot_wider(names_from = funding_year, names_prefix = "Commitment_", values_from = tot) %>% 
  mutate_if(is.numeric, .funs = list(perc = ~ (function(x) {x / sum(x)})(.))) %>% # https://stackoverflow.com/a/57171049/5593458
  relocate(Category, Commitment_2016, Commitment_2016_perc, Commitment_2017, Commitment_2017_perc, Commitment_2018, Commitment_2018_perc, Commitment_2019, Commitment_2019_perc, Commitment_2020, Commitment_2020_perc) %>% 
  janitor::adorn_totals("row") %>%  # https://stackoverflow.com/a/49819006/5593458
  gt() %>% #https://themockup.blog/posts/2020-09-04-10-table-rules-in-r/
  # https://themockup.blog/posts/2020-05-16-gt-a-grammer-of-tables/
  fmt_currency(
    columns = vars(Commitment_2016, Commitment_2017, Commitment_2018, Commitment_2019, Commitment_2020),
    currency = "USD"
  ) %>% 
  fmt_percent(
    columns = vars(Commitment_2016_perc, Commitment_2017_perc, Commitment_2018_perc, Commitment_2019_perc, Commitment_2020_perc),
    scale_values = T
  ) %>%
  cols_align(align = "left",
             columns = 'Category') %>%
  cols_align(align = "right",
             columns = vars(Commitment_2016, Commitment_2016_perc, Commitment_2017, Commitment_2017_perc, Commitment_2018, Commitment_2018_perc, Commitment_2019, Commitment_2019_perc, Commitment_2020, Commitment_2020_perc)) %>% 
  cols_label(
    Category = "",
    Commitment_2016 = "Yearly Total",
    Commitment_2016_perc = "%",
    Commitment_2017 = "Yearly Total", 
    Commitment_2017_perc = "%",
    Commitment_2018 = "Yearly Total", 
    Commitment_2018_perc = "%", 
    Commitment_2019 = "Yearly Total", 
    Commitment_2019_perc = "%", 
    Commitment_2020 = "Yearly Total", 
    Commitment_2020_perc = "%"
  ) %>% 
  tab_header(
    title = md("**National Summary of Library Entity Commitments**")
    ) %>% 
  tab_spanner(
    label = "2016",
    columns = c(2:3)
  ) %>% 
  tab_spanner(
    label = "2017",
    columns = c(4:5)
  ) %>% 
  tab_spanner(
    label = "2018",
    columns = c(6:7)
  ) %>% 
  tab_spanner(
    label = "2019",
    columns = c(8:9)
  ) %>% 
  tab_spanner(
    label = "2020",
    columns = c(10:11)
  ) %>% 
   tab_style(
    style = cell_fill(
      color = "#E8E8E8"
    ),
    locations = list(
      cells_body(
        columns = vars(Commitment_2016_perc, Commitment_2017_perc, Commitment_2018_perc, Commitment_2019_perc, Commitment_2020_perc)
        ),
      cells_column_labels(
        columns = vars(Commitment_2016_perc, Commitment_2017_perc, Commitment_2018_perc, Commitment_2019_perc, Commitment_2020_perc)
      )
      )
  ) %>%
  tab_style(
    style = cell_borders(
      sides = "top",
      color = "black",
      weight = px(2)
    ),
    locations = cells_body(
      columns = everything(),
      rows = Category == "Total"
    )
  ) %>% 
  opt_table_lines(extent = "default") %>%
  tab_options(
      heading.align = "left",
      row_group.border.top.width = px(3),
      row_group.border.top.color = "black",
      row_group.border.bottom.color = "black",
      table_body.hlines.color = "white",
      table.border.top.color = "#3778bf",
      table.border.top.width = px(3),
      table.border.bottom.color = "#3778bf",
      table.border.bottom.width = px(3),
      column_labels.border.bottom.color = "black",
      column_labels.border.bottom.width = px(2),
      heading.border.bottom.color = "#3778bf",
      heading.border.bottom.width = px(3)
    ) 
```


```{r}
erate_libs %>% 
  select(ros_entity_number, funding_year, chosen_category_of_service, cat2_discount_by_ros, cat1_discount_by_ros_estimated) %>%
  group_by(funding_year) %>%
  summarise(sum_cat1_discount = sum(cat1_discount_by_ros_estimated, na.rm=T),
            sum_cat2_allocation = sum(cat2_discount_by_ros, na.rm = T),
            nat_total = sum_cat1_discount + sum_cat2_allocation,
            total_entities = n_distinct(ros_entity_number),
            per_entity_commitment = nat_total/total_entities) %>% 
  select(funding_year, nat_total, total_entities, per_entity_commitment) 
```


```{r}
# Visualize above data
# Color palette: "#3778bf", "#feb209", "#b0054b"
erate_libs %>%
  select(funding_year,
         chosen_category_of_service,
         cat2_discount_by_ros,
         cat1_discount_by_ros_estimated) %>%
  group_by(funding_year) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(Category1_Commitment = if(all(is.na(cat1_discount_by_ros_estimated))) NA_real_ 
            else round(sum(cat1_discount_by_ros_estimated, na.rm = T), 2),
            Category2_Commitment = if(all(is.na(cat2_discount_by_ros))) NA_real_ 
            else round(sum(cat2_discount_by_ros, na.rm = T), 2)) %>%
  rowwise() %>% 
  mutate(National_Total = sum(c(Category1_Commitment, Category2_Commitment), na.rm = TRUE)) %>%
  pivot_longer(c(Category1_Commitment, Category2_Commitment), names_to = "Cat") %>% 
  mutate(Pct_label = round((value/National_Total)*100, 1)) %>% 
  ggplot(aes(x = funding_year, y = value, fill = Cat)) +
  geom_bar(position="fill", stat="identity", color="black") +
  scale_fill_manual(values=c("#3778bf", "#feb209"), name = "Funding Category", labels = c("Cat 1", "Cat 2")) +
  geom_label(aes(label=paste0(scales::dollar(value), "\n", Pct_label,"%")), color="black",
            position = position_fill(vjust = 0.45), size=2.5, show.legend = FALSE) +
  theme_linedraw() +
  theme(axis.title.y=element_blank(),
        axis.title.x=element_blank()) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(title = "Library Commitment Breakdown by Year")
```


```{r}
# Create a table with national Cat1 and Cat2 totals by year broken out by urban rural status
erate_libs %>%
  select(
    funding_year,
    chosen_category_of_service,
    ros_urban_rural_status,
    cat2_discount_by_ros,
    cat1_discount_by_ros_estimated
  ) %>%
  group_by(funding_year, ros_urban_rural_status) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(Category1_Commitment = if(all(is.na(cat1_discount_by_ros_estimated))) NA_real_ 
            else sum(cat1_discount_by_ros_estimated, na.rm = T),
            Category2_Commitment = if(all(is.na(cat2_discount_by_ros))) NA_real_ 
            else sum(cat2_discount_by_ros, na.rm = T)) %>%
  rowwise() %>% 
  mutate(National_Total = sum(c(Category1_Commitment, Category2_Commitment), na.rm = TRUE)) %>% 
  gt(groupname_col = "funding_year") %>% #https://themockup.blog/posts/2020-09-04-10-table-rules-in-r/
  # https://themockup.blog/posts/2020-05-16-gt-a-grammer-of-tables/
  fmt_currency(
    columns = vars(Category1_Commitment, Category2_Commitment, National_Total), # reference cols by name
    currency = "USD"
  ) %>% 
  fmt_missing(
    columns = vars(ros_urban_rural_status),
    missing_text = "Unassigned"
  ) %>%
  fmt_missing(
    columns = vars(Category1_Commitment, Category2_Commitment),
    missing_text = "Not Applicable"
  ) %>% 
  cols_align(align = "left",
             columns = 'funding_year') %>%
  cols_align(align = "center",
             columns = 'ros_urban_rural_status') %>%
  cols_align(align = "right",
             columns = vars(Category1_Commitment, Category2_Commitment, National_Total)) %>% 
  cols_width(vars(ros_urban_rural_status) ~ px(90),
             vars(Category1_Commitment, Category2_Commitment, National_Total) ~ px(170)
  ) %>% 
  cols_label(
    funding_year = "",
    ros_urban_rural_status = md("Geography"),
    Category1_Commitment = md("Category 1<br>Commitment"),
    Category2_Commitment = md("Category 2<br>Commitment"),
    National_Total = md("Total<br>Commitment")
  ) %>% 
  tab_header(
    title = md("**National Summary of Library Entities by Geography**")
    ) %>% 
  tab_style(
    style = cell_fill(
      color = "#DCDCDC"
    ),
    locations = list(
      cells_body(
        columns = vars(National_Total)
        )
      )
      ) %>%
  tab_style(
    style = cell_text(color = "black", weight = "bold"),
    locations = list(
      cells_row_groups(),
      cells_column_labels(everything())
    )
  ) %>% 
  tab_style(
    style = list(
      cell_borders(
        sides = "left",
        color = "black",
        weight = px(3)
      )
    ),
    locations = list(
      cells_body(
        columns = vars(National_Total)
      )
    )
    ) %>% 
  opt_table_lines(extent = "default") %>%
  tab_options(
    row_group.border.top.width = px(3),
    row_group.border.top.color = "black",
    row_group.border.bottom.color = "black",
    table_body.hlines.color = "white",
    table.border.top.color = "#3778bf",
    table.border.top.width = px(3),
    table.border.bottom.color = "#3778bf",
    table.border.bottom.width = px(3),
    column_labels.border.bottom.color = "black",
    column_labels.border.bottom.width = px(2),
    heading.border.bottom.color = "#3778bf",
    heading.border.bottom.width = px(3)
  ) 
```

```{r}
# Create a table with national Cat1 and Cat2 totals by year
erate_libs %>%
  mutate(Commitment = case_when(
    is.na(cat1_discount_by_ros_estimated) ~ cat2_discount_by_ros,
    is.na(cat2_discount_by_ros) ~ cat1_discount_by_ros_estimated
  )) %>% 
  select(
    funding_year,
    chosen_category_of_service,
    ros_urban_rural_status,
    Commitment
  ) %>%
  group_by(funding_year,
           chosen_category_of_service,
           ros_urban_rural_status) %>%
  summarise(Commitment = sum(Commitment, na.rm = T)) %>%
  ungroup() %>% 
  group_by(funding_year) %>% 
  mutate(Pct = Commitment / sum(Commitment)) %>% 
  pivot_wider(names_from = funding_year, values_from = c(Commitment, Pct), names_sort = T) %>% 
  relocate(chosen_category_of_service, ros_urban_rural_status, Commitment_2016, Pct_2016, Commitment_2017, Pct_2017, Commitment_2018, Pct_2018, Commitment_2019, Pct_2019, Commitment_2020, Pct_2020) %>% 
  janitor::adorn_totals("row") %>%  # https://stackoverflow.com/a/49819006/5593458
  ungroup() %>% 
  mutate(ros_urban_rural_status = replace_na(ros_urban_rural_status, "Unassigned")) %>% 
  gt(groupname_col = "ros_urban_rural_status",
     rowname_col = "chosen_category_of_service") %>% #https://themockup.blog/posts/2020-09-04-10-table-rules-in-r/
  # https://themockup.blog/posts/2020-05-16-gt-a-grammer-of-tables/
  fmt_currency(
    columns = vars(Commitment_2016, Commitment_2017, Commitment_2018, Commitment_2019, Commitment_2020),
    currency = "USD"
  ) %>% 
  fmt_percent(
    columns = vars(Pct_2016, Pct_2017, Pct_2018, Pct_2019, Pct_2020),
    scale_values = T
  ) %>%
  fmt_missing(
    columns = vars(Commitment_2016, Pct_2016, Commitment_2017, Pct_2017, Commitment_2018, Pct_2018, Commitment_2019, Pct_2019, Commitment_2020, Pct_2020),
    missing_text = "None"
  ) %>%
  cols_align(align = "left",
             columns = 'ros_urban_rural_status') %>%
  cols_align(align = "center",
             columns = 'chosen_category_of_service') %>%
  cols_align(align = "right",
             columns = vars(Commitment_2016, Pct_2016, Commitment_2017, Pct_2017, Commitment_2018, Pct_2018, Commitment_2019, Pct_2019, Commitment_2020, Pct_2020)) %>% 
  cols_label(
    Commitment_2016 = "Yearly Total",
    Pct_2016 = "%",
    Commitment_2017 = "Yearly Total", 
    Pct_2017 = "%",
    Commitment_2018 = "Yearly Total", 
    Pct_2018 = "%", 
    Commitment_2019 = "Yearly Total", 
    Pct_2019 = "%", 
    Commitment_2020 = "Yearly Total", 
    Pct_2020 = "%"
  ) %>% 
  tab_header(
    title = md("**National Summary of Urban and Rural Commitments by Category**")
    ) %>% 
  tab_spanner(
    label = "2016",
    columns = c(3:4)
  ) %>% 
  tab_spanner(
    label = "2017",
    columns = c(5:6)
  ) %>% 
  tab_spanner(
    label = "2018",
    columns = c(7:8)
  ) %>% 
  tab_spanner(
    label = "2019",
    columns = c(9:10)
  ) %>% 
  tab_spanner(
    label = "2020",
    columns = c(11:12)
  ) %>% 
   tab_style(
    style = cell_fill(
      color = "#E8E8E8"
    ),
    locations = list(
      cells_body(
        columns = vars(Pct_2016, Pct_2017, Pct_2018, Pct_2019, Pct_2020)
        ),
      cells_column_labels(
        columns = vars(Pct_2016, Pct_2017, Pct_2018, Pct_2019, Pct_2020)
      )
      )
  ) %>%
  tab_style(
    style = cell_text(color = "black", weight = "bold"),
    locations = list(
      cells_row_groups(),
      cells_column_labels(everything()),
      cells_column_spanners(everything())
    )
  ) %>% 
  tab_style(
    style = cell_borders(
      sides = "top",
      color = "black",
      weight = px(2)
    ),
    locations = cells_body(
      columns = everything(),
      rows = ros_urban_rural_status == "Total"
    )
  ) %>% 
  tab_options(
    heading.align = "left",
    row_group.border.top.width = px(3),
    row_group.border.top.color = "black",
    row_group.border.bottom.color = "black",
    table_body.hlines.color = "white",
    table.border.top.color = "#3778bf",
    table.border.top.width = px(3),
    table.border.bottom.color = "#3778bf",
    table.border.bottom.width = px(3),
    column_labels.border.bottom.color = "black",
    column_labels.border.bottom.width = px(2),
    heading.border.bottom.color = "#3778bf",
    heading.border.bottom.width = px(3)
  ) 
```


```{r}
# Create a table with national Cat1 and Cat2 totals by year for IMLS recipients only
erate_imls %>%
  filter(!is.na(FSCSKEY)) %>%
  select(funding_year,
         chosen_category_of_service,
         cat2_discount_by_ros,
         cat1_discount_by_ros_estimated) %>%
  group_by(funding_year) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(Category1_Commitment = if(all(is.na(cat1_discount_by_ros_estimated))) NA_real_ 
            else sum(cat1_discount_by_ros_estimated, na.rm = T),
            Category2_Commitment = if(all(is.na(cat2_discount_by_ros))) NA_real_ 
            else sum(cat2_discount_by_ros, na.rm = T)) %>%
  rowwise() %>% 
  mutate(National_Total = sum(c(Category1_Commitment, Category2_Commitment), na.rm = TRUE))
```

```{r}
# Create a table with national Cat1 and Cat2 totals by year for IMLS recipients only broken out
# into USAC urban rural status
erate_imls %>%
  filter(!is.na(FSCSKEY)) %>%
  select(
    funding_year,
    chosen_category_of_service,
    ros_urban_rural_status,
    cat2_discount_by_ros,
    cat1_discount_by_ros_estimated
  ) %>%
  group_by(funding_year, ros_urban_rural_status) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(Category1_Commitment = if(all(is.na(cat1_discount_by_ros_estimated))) NA_real_ 
            else sum(cat1_discount_by_ros_estimated, na.rm = T),
            Category2_Commitment = if(all(is.na(cat2_discount_by_ros))) NA_real_ 
            else sum(cat2_discount_by_ros, na.rm = T)) %>%
  rowwise() %>% 
  mutate(National_Total = sum(c(Category1_Commitment, Category2_Commitment), na.rm = TRUE))
```

```{r}
# Create a table with national Cat1 and Cat2 totals by year for IMLS Outlets only
imls_outlet_erate %>%
  select(funding_year,
         chosen_category_of_service,
         cat2_discount_by_ros,
         cat1_discount_by_ros_estimated) %>%
  group_by(funding_year) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(Category1_Commitment = if(all(is.na(cat1_discount_by_ros_estimated))) NA_real_ 
            else sum(cat1_discount_by_ros_estimated, na.rm = T),
            Category2_Commitment = if(all(is.na(cat2_discount_by_ros))) NA_real_ 
            else sum(cat2_discount_by_ros, na.rm = T)) %>%
  rowwise() %>% 
  mutate(National_Total = sum(c(Category1_Commitment, Category2_Commitment), na.rm = TRUE))
```

```{r}
# Create a table with national Cat1 and Cat2 totals by year by urban rural status
# for imls outlets only
imls_outlet_erate %>%
  select(
    funding_year,
    chosen_category_of_service,
    ros_urban_rural_status,
    cat2_discount_by_ros,
    cat1_discount_by_ros_estimated
  ) %>%
  group_by(funding_year, ros_urban_rural_status) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(
    Category1_Commitment = if (all(is.na(cat1_discount_by_ros_estimated)))
      NA_real_
    else
      sum(cat1_discount_by_ros_estimated, na.rm = T),
    Category2_Commitment = if (all(is.na(cat2_discount_by_ros)))
      NA_real_
    else
      sum(cat2_discount_by_ros, na.rm = T)
  ) %>%
  rowwise() %>%
  mutate(National_Total = sum(c(
    Category1_Commitment, Category2_Commitment
  ), na.rm = TRUE))
```

```{r}
# Find out which outlets match to more than one REN (ros entity number)
imls_outlet_erate %>% 
  group_by(FSCSKEY, FSCS_SEQ, .drop = F) %>%  
  summarise(CountRENS = n_distinct(ros_entity_number)) %>% 
  filter(CountRENS > 1)
```

```{r}
# assign the total number of outlets on the dataset to a variable
total_imls <- (ae_and_outlets %>% summarise(n_distinct(FSCSKEY, FSCS_SEQ)))[[1]]
```

```{r}
# What percentage of outlets applied for erate funding in 2019?
erate_imls %>% 
  filter(!(is.na(FSCSKEY) & is.na(FSCS_SEQ))) %>%
  group_by(FSCSKEY, FSCS_SEQ, funding_year, .drop = F) %>% 
  # To adjust for the fact that some FSCSs have multiple RENs, 
  # we'll just take the first one that shows up
  slice_head(n = 1) %>% 
  ungroup() %>% 
  group_by(funding_year) %>% 
  summarise(countRENS = n_distinct(ros_entity_number, na.rm = T),
           Pct = n_distinct(ros_entity_number, na.rm = T)/total_imls)
```

```{r}
erate_imls %>% 
  filter(!(is.na(FSCSKEY) & is.na(FSCS_SEQ))) %>%
  group_by(FSCSKEY, FSCS_SEQ, funding_year, .drop = F) %>% 
  # To adjust for the fact that some FSCSs have multiple RENs, 
  # we'll just take the first one that shows up
  slice_head(n = 1) %>% 
  ungroup() %>% 
  group_by(funding_year) %>% 
  summarise(countRENS = n_distinct(ros_entity_number, na.rm = T),
           Pct = n_distinct(ros_entity_number, na.rm = T)/total_imls) %>% 
  ggplot(aes(x = funding_year, y = Pct, label = glue("{round(Pct*100, 1)}%"))) +
  geom_line(color = "#3778bf") +
  geom_point(color = "#b0054b") +
  geom_label(vjust = 0, nudge_y = 0.06) +
  theme_linedraw() +
  #scale_color_manual(values = c("#3778bf", "#feb209", "#b0054b")) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1)) +
  labs(title = "E-rate Participation by IMLS Registered Entities",
       x = "Funding Year",
       y = "Pct of Entities Participating",
       caption = "Total number of IMLS Entities (i.e. the denominator) = 17635\n A Combination of Admin Entities and Outlets")

#ggsave("~/Google Drive/Shared drives/ALA-ERate/Plots/IMLS_Library_Participation_linechart.png")
```


```{r}
# Chris/Marijke request
# Chart 1: Percentage of participation per state of IMLS outlets by Category of Service
# Category 1 only - Category 2 only - Category 1 & 2

# imls_outlet_erate %>%
#   filter(funding_year == "2019") %>%
#   group_by(FSCSKEY, FSCS_SEQ, chosen_category_of_service, .drop = F) %>%
#   slice_head(n = 1) %>%
#   ungroup() %>%
#   group_by(STABR, chosen_category_of_service, .drop = F) %>%
#   summarise(Cnt_of_Outlets_Recving_Erate = n_distinct(ros_entity_number, na.rm = T)) %>%
#   bind_rows(imls_outlet_erate %>%
#               filter(funding_year == "2019") %>%
#               group_by(FSCSKEY, FSCS_SEQ, .drop = F) %>%
#               mutate(How_Many_Categories = n_distinct(chosen_category_of_service)) %>% 
#               filter(How_Many_Categories > 1) %>% 
#               slice_head(n = 1) %>% 
#               group_by(STABR) %>% 
#               tally(name = "Cnt_of_Outlets_Recving_Erate") %>% 
#               mutate(chosen_category_of_service = "Category1AND2")) %>% 
#   left_join(imls_outlet_erate %>%
#               group_by(STABR, .drop = F) %>%
#               summarise(Total_Cnt_Outlets = n_distinct(FSCSKEY, FSCS_SEQ))) %>% 
#   mutate(Pct_Outlets_Recving_Erate = paste0(round((Cnt_of_Outlets_Recving_Erate / Total_Cnt_Outlets) * 100, 2), "%")) %>%
#   filter(!is.na(chosen_category_of_service)) %>%
#   select(STABR, chosen_category_of_service, Pct_Outlets_Recving_Erate) %>% 
#   pivot_wider(names_from = chosen_category_of_service, values_from = Pct_Outlets_Recving_Erate) %>%
#   write.csv(paste0("~/Google Drive File Stream/Shared drives/ALA-ERate/Data/2019_Pct_Participation_by_State_by_Category_", Sys.Date(), ".csv"))

imls_outlet_erate %>%
  filter(funding_year == "2019") %>%
  group_by(FSCSKEY, FSCS_SEQ, .drop = F) %>%
  mutate(How_Many_Categories = n_distinct(chosen_category_of_service)) %>% 
  filter(How_Many_Categories == 1) %>% 
  slice_head(n = 1) %>% 
  group_by(STABR, chosen_category_of_service) %>% 
  tally(name = "Cnt_of_Outlets_Recving_Erate") %>% 
  bind_rows(imls_outlet_erate %>%
              filter(funding_year == "2019") %>%
              group_by(FSCSKEY, FSCS_SEQ, .drop = F) %>%
              mutate(How_Many_Categories = n_distinct(chosen_category_of_service)) %>% 
              filter(How_Many_Categories > 1) %>% 
              slice_head(n = 1) %>% 
              group_by(STABR) %>% 
              tally(name = "Cnt_of_Outlets_Recving_Erate") %>% 
              mutate(chosen_category_of_service = "Category1AND2")) %>% 
  left_join(imls_outlet_erate %>%
              group_by(STABR, .drop = F) %>%
              summarise(Total_Cnt_Outlets = n_distinct(FSCSKEY, FSCS_SEQ))) %>% 
  mutate(Pct_Outlets_Recving_Erate = paste0(round((Cnt_of_Outlets_Recving_Erate / Total_Cnt_Outlets) * 100, 2), "%")) %>% 
  select(STABR, chosen_category_of_service, Cnt_of_Outlets_Recving_Erate, Pct_Outlets_Recving_Erate) %>% 
  pivot_wider(names_from = chosen_category_of_service, values_from = c(Cnt_of_Outlets_Recving_Erate, Pct_Outlets_Recving_Erate)) #%>% 
  #write.csv(paste0("~/Google Drive File Stream/Shared drives/ALA-ERate/Data/2019_Pct_Participation_by_State_by_Category_", Sys.Date(), ".csv"))

```

```{r}
imls_outlet_erate %>%
  filter(funding_year == "2019") %>%
  group_by(FSCSKEY, FSCS_SEQ, .drop = F) %>%
  mutate(How_Many_Categories = n_distinct(chosen_category_of_service)) %>% 
  filter(How_Many_Categories > 1) %>% 
  slice_head(n = 1) %>% 
  group_by(STABR) %>% 
  tally(name = "Cnt_of_Outlets_Recving_Erate") %>% 
  mutate(chosen_category_of_service = "Category1AND2")
```

```{r}
# Chris/Marijke request
# Chart 1: Percentage of participation per state of IMLS outlets by Category of Service
# Category 1 only - Category 2 only - Category 1 & 2
imls_outlet_erate %>%
  filter(funding_year == "2019") %>%
  group_by(FSCSKEY, FSCS_SEQ, .drop = F) %>%
  slice_head(n = 1) %>%
  ungroup() %>%
  group_by(STABR, .drop = F) %>%
  summarise(Cnt_of_Outlets_Recving_Erate = n_distinct(ros_entity_number, na.rm = T)) %>%
  left_join(imls_outlet_erate %>%
              group_by(STABR, .drop = F) %>%
              summarise(Total_Cnt_Outlets = n_distinct(FSCSKEY, FSCS_SEQ))) %>%
  mutate(PctParticipating = paste0(round((Cnt_of_Outlets_Recving_Erate / Total_Cnt_Outlets) * 100, 2), "%")) #%>% 
  #write.csv(paste0("~/Google Drive File Stream/Shared drives/ALA-ERate/Data/2019_Pct_Participation_by_State_", Sys.Date(), ".csv"))
```

```{r}
# Chris/Marijke request
# Chart 2: Percentage of Participation per state of IMLS outlets by Category of Service
# Consortium/  Library / Library System
# Category 1 only - Category 2 only - Category 1 & 2

imls_outlet_erate %>%
  filter(funding_year == "2019") %>%
  group_by(FSCSKEY, FSCS_SEQ, ros_entity_number, chosen_category_of_service, .drop = F) %>%
  slice_head(n = 1) %>% 
  ungroup() %>%
  group_by(STABR, chosen_category_of_service, .drop = F) %>%
  summarise(n_distinct(ros_entity_number, na.rm = T)) #%>% write.csv(paste0("~/Google Drive File Stream/Shared drives/ALA-ERate/Data/2019_Unique_Entities_by_State_by_Category_", Sys.Date(), ".csv"))
```

```{r}
# Chris/Marijke request
# Chart 2: Percentage of Participation per state of IMLS outlets by Category of Service
# Consortium/  Library / Library System
# Category 1 only - Category 2 only - Category 1 & 2

imls_outlet_erate %>%
  filter(funding_year == "2019") %>%
  group_by(FSCSKEY, FSCS_SEQ, chosen_category_of_service, form_471_service_type_name, form_471_function_name, form_471_product_name, .drop = F) %>%
  slice_head(n = 1) %>% 
  ungroup() %>%
  group_by(chosen_category_of_service, form_471_service_type_name, form_471_function_name, form_471_product_name, .drop = F) %>%
  summarise(n_distinct(ros_entity_number, na.rm = T)) #%>% 
  #write.csv(paste0("~/Google Drive File Stream/Shared drives/ALA-ERate/Data/2019_Service_Products_IMLS_Outlets_", Sys.Date(), ".csv"))
```

```{r}
# Chris/Marijke request
# Chart 2: Percentage of Participation per state of IMLS outlets by Category of Service
# Consortium/  Library / Library System
# Category 1 only - Category 2 only - Category 1 & 2

imls_outlet_erate %>%
  group_by(FSCSKEY, FSCS_SEQ, chosen_category_of_service, form_471_service_type_name, form_471_function_name, form_471_product_name, .drop = F) %>%
  slice_head(n = 1) %>% 
  ungroup() %>%
  group_by(chosen_category_of_service, form_471_service_type_name, form_471_function_name, form_471_product_name, funding_year, .drop = F) %>%
  summarise(n_distinct(ros_entity_number, na.rm = T)) #%>% 
  #write.csv(paste0("~/Google Drive File Stream/Shared drives/ALA-ERate/Data/AllYears_Service_Products_IMLS_Outlets_", Sys.Date(), ".csv"))
```

```{r}
# imls_outlet_erate %>%
#   filter(funding_year == "2019") %>%
#   group_by(FSCSKEY, FSCS_SEQ, chosen_category_of_service, .drop = F) %>%
#   slice_head(n = 1) %>%
#   ungroup() %>%
#   group_by(STABR, chosen_category_of_service, .drop = F) %>%
#   summarise(Cnt_of_Outlets_Recving_Erate = n_distinct(ros_entity_number, na.rm = T)) %>%
#   bind_rows(imls_outlet_erate %>%
#               filter(funding_year == "2019") %>%
#               group_by(FSCSKEY, FSCS_SEQ, .drop = F) %>%
#               mutate(How_Many_Categories = n_distinct(chosen_category_of_service)) %>% 
#               filter(How_Many_Categories > 1) %>% 
#               slice_head(n = 1) %>% 
#               group_by(STABR, .drop = F) %>% 
#               tally(name = "Cnt_of_Outlets_Recving_Erate") %>% 
#               mutate(chosen_category_of_service = "Category1AND2")) %>% 
#   left_join(imls_outlet_erate %>%
#               group_by(STABR, .drop = F) %>%
#               summarise(Total_Cnt_Outlets = n_distinct(FSCSKEY, FSCS_SEQ))) %>% 
#   mutate(Pct_Outlets_Recving_Erate = paste0(round((Cnt_of_Outlets_Recving_Erate / Total_Cnt_Outlets) * 100, 2), "%")) %>%
#   filter(!is.na(chosen_category_of_service))
```

```{r}
# Chris/Marijke request
# Chart 2: Percentage of Participation per state of IMLS outlets by Category of Service
# Consortium/  Library / Library System
# Category 1 only - Category 2 only - Category 1 & 2

imls_outlet_erate %>%
  filter(funding_year == "2019") %>%
  group_by(FSCSKEY, FSCS_SEQ, chosen_category_of_service, .drop = F) %>%
  slice_head(n = 1) %>%
  ungroup() %>%
  group_by(STABR, chosen_category_of_service, .drop = F) %>%
  mutate(Cnt_of_Outlets_Recving_Erate_by_Cat = n_distinct(ros_entity_number, na.rm = T)) %>% 
  ungroup() %>%
  group_by(STABR,
           chosen_category_of_service,
           organization_entity_type_name,
           Cnt_of_Outlets_Recving_Erate_by_Cat,
           .drop = F) %>%
  summarise(Cnt_of_Outlets_Recving_Erate_by_Org = n_distinct(ros_entity_number, na.rm = T)) %>%
  mutate(PctApplyingAs = paste0(round((Cnt_of_Outlets_Recving_Erate_by_Org / Cnt_of_Outlets_Recving_Erate_by_Cat) * 100, 2), "%")) #%>% 
  #write.csv(paste0("~/Google Drive File Stream/Shared drives/ALA-ERate/Data/2019_Pct_Participation_by_State_by_Category_by_OrgType_", Sys.Date(), ".csv"))
```

```{r}
# Chris/Marijke request
# Chart 2: Percentage of Participation per state of IMLS outlets by Category of Service
# Consortium/  Library / Library System
# Category 1 only - Category 2 only - Category 1 & 2

imls_outlet_erate %>%
  filter(funding_year == "2019") %>%
  group_by(FSCSKEY, FSCS_SEQ, chosen_category_of_service, .drop = F) %>%
  slice_head(n = 1) %>%
  ungroup() %>%
  group_by(STABR, chosen_category_of_service, .drop = F) %>%
  mutate(Cnt_of_Outlets_Recving_Erate_by_Cat = n_distinct(ros_entity_number, na.rm = T)) %>% 
  ungroup() %>%
  group_by(STABR,
           chosen_category_of_service,
           organization_entity_type_name,
           Cnt_of_Outlets_Recving_Erate_by_Cat,
           .drop = F) %>%
  summarise(Cnt_of_Outlets_Recving_Erate_by_Org = n_distinct(ros_entity_number, na.rm = T)) %>%
  mutate(PctApplyingAs = Cnt_of_Outlets_Recving_Erate_by_Org / Cnt_of_Outlets_Recving_Erate_by_Cat) %>%
  ggplot(aes(x = chosen_category_of_service, y = PctApplyingAs)) +
  # preserve = single makes each bar the same width
  geom_bar(aes(fill = organization_entity_type_name), 
           position = position_dodge(preserve = "single"), 
           stat = "identity",
           color="black") +
  theme_linedraw() +
  labs(title = "How IMLS Outlets Applied for Erate in 2019",
       subtitle = "by State and Category of Service",
       y = "Percentage of Total Applying \nWithin State and Category", 
       x = "Category of Service",
       fill = "Org. Type") +
  scale_y_continuous(breaks=c(0, .25, .50, .75, 1), limits = c(0, 1), labels = scales::percent) +
  scale_fill_manual(values = c("#3778bf", "#feb209", "#b0054b")) +
  facet_wrap(~ STABR, ncol = 5, drop = FALSE) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

ggsave(paste0("~/Google Drive File Stream/Shared drives/ALA-ERate/Plots/2019_Pct_Participation_by_State_by_Category_by_OrgType_", Sys.Date(), ".pdf"), width = 8, height = 16.5, units = "in")
```

```{r}
# Create a table to show which service type is associated with which category
erate_libs %>%
  distinct(chosen_category_of_service, form_471_service_type_name)
```

```{r}
# Create a table with national Cat1 and Cat2 totals by year, service type, and function
erate_libs %>%
  select(
    funding_year,
    chosen_category_of_service,
    form_471_service_type_name,
    form_471_function_name,
    cat2_discount_by_ros,
    cat1_discount_by_ros_estimated
  ) %>%
  group_by(funding_year,
           form_471_service_type_name,
           form_471_function_name) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(
    Category1_Commitment = if (all(is.na(cat1_discount_by_ros_estimated)))
      NA_real_
    else
      sum(cat1_discount_by_ros_estimated, na.rm = T),
    Category2_Commitment = if (all(is.na(cat2_discount_by_ros)))
      NA_real_
    else
      sum(cat2_discount_by_ros, na.rm = T)
  ) %>%
  rowwise() %>%
  mutate(National_Total = sum(c(
    Category1_Commitment, Category2_Commitment
  ), na.rm = TRUE))
```

```{r}
# Category 1 service type & function totals over time
erate_libs %>%
  filter(chosen_category_of_service == "Category1") %>% 
  select(funding_year, form_471_service_type_name, form_471_function_name,  cat2_discount_by_ros, cat1_discount_by_ros_estimated) %>%
  group_by(funding_year, form_471_service_type_name, form_471_function_name) %>%
  summarise(Commitment = sum(cat1_discount_by_ros_estimated, na.rm=T)) %>% 
  pivot_wider(names_from = funding_year, values_from = Commitment)
```

```{r}
# Category 2 service type & function totals over time
erate_libs %>%
  filter(chosen_category_of_service == "Category2") %>% 
  select(funding_year, form_471_service_type_name, form_471_function_name,  cat2_discount_by_ros, cat1_discount_by_ros_estimated) %>%
  group_by(funding_year, form_471_service_type_name, form_471_function_name) %>%
  summarise(Commitment = sum(cat2_discount_by_ros, na.rm = T)) %>% 
  pivot_wider(names_from = funding_year, values_from = Commitment)
```

```{r}
# Create a table with national Cat1 and Cat2 totals by year
erate_libs %>%
  mutate(Commitment = case_when(
    is.na(cat1_discount_by_ros_estimated) ~ cat2_discount_by_ros,
    is.na(cat2_discount_by_ros) ~ cat1_discount_by_ros_estimated
  )) %>% 
  select(
    funding_year,
    form_471_service_type_name,
    Commitment
  ) %>%
  group_by(funding_year,
           form_471_service_type_name) %>%
  summarise(Commitment = sum(Commitment, na.rm = T)) %>%
  mutate(Pct = Commitment / sum(Commitment)) %>% 
  pivot_wider(names_from = funding_year, values_from = c(Commitment, Pct), names_sort = T) %>% 
  relocate(form_471_service_type_name, Commitment_2016, Pct_2016, Commitment_2017, Pct_2017, Commitment_2018, Pct_2018, Commitment_2019, Pct_2019, Commitment_2020, Pct_2020) %>% 
  arrange(desc(Pct_2016)) %>% 
  janitor::adorn_totals("row") %>%  # https://stackoverflow.com/a/49819006/5593458
  gt() %>% #https://themockup.blog/posts/2020-09-04-10-table-rules-in-r/
  # https://themockup.blog/posts/2020-05-16-gt-a-grammer-of-tables/
  fmt_currency(
    columns = vars(Commitment_2016, Commitment_2017, Commitment_2018, Commitment_2019, Commitment_2020),
    currency = "USD"
  ) %>% 
  fmt_percent(
    columns = vars(Pct_2016, Pct_2017, Pct_2018, Pct_2019, Pct_2020),
    scale_values = T
  ) %>%
  cols_align(align = "left",
             columns = 'form_471_service_type_name') %>%
  cols_align(align = "right",
             columns = vars(Commitment_2016, Pct_2016, Commitment_2017, Pct_2017, Commitment_2018, Pct_2018, Commitment_2019, Pct_2019, Commitment_2020, Pct_2020)) %>% 
  # cols_width(vars(ros_urban_rural_status) ~ px(90),
  #            vars(Category1_Commitment, Category2_Commitment, National_Total) ~ px(170)
  # ) %>% 
  cols_label(
    form_471_service_type_name = "Service Type",
    Commitment_2016 = "Yearly Total",
    Pct_2016 = "%",
    Commitment_2017 = "Yearly Total", 
    Pct_2017 = "%",
    Commitment_2018 = "Yearly Total", 
    Pct_2018 = "%", 
    Commitment_2019 = "Yearly Total", 
    Pct_2019 = "%", 
    Commitment_2020 = "Yearly Total", 
    Pct_2020 = "%"
  ) %>% 
  tab_header(
    title = md("**National Summary of Service Type Commitments**")
    ) %>% 
  tab_spanner(
    label = "2016",
    columns = c(2:3)
  ) %>% 
  tab_spanner(
    label = "2017",
    columns = c(4:5)
  ) %>% 
  tab_spanner(
    label = "2018",
    columns = c(6:7)
  ) %>% 
  tab_spanner(
    label = "2019",
    columns = c(8:9)
  ) %>% 
  tab_spanner(
    label = "2020",
    columns = c(10:11)
  ) %>% 
   tab_style(
    style = cell_fill(
      color = "#E8E8E8"
    ),
    locations = list(
      cells_body(
        columns = vars(Pct_2016, Pct_2017, Pct_2018, Pct_2019, Pct_2020)
        ),
      cells_column_labels(
        columns = vars(Pct_2016, Pct_2017, Pct_2018, Pct_2019, Pct_2020)
      )
      )
  ) %>%
  tab_style(
    style = cell_borders(
      sides = "top",
      color = "black",
      weight = px(2)
    ),
    locations = cells_body(
      columns = everything(),
      rows = form_471_service_type_name == "Total"
    )
  ) %>% 
  opt_table_lines(extent = "default") %>%
  tab_options(
    heading.align = "left",
    row_group.border.top.width = px(3),
    row_group.border.top.color = "black",
    row_group.border.bottom.color = "black",
    table_body.hlines.color = "white",
    table.border.top.color = "#3778bf",
    table.border.top.width = px(3),
    table.border.bottom.color = "#3778bf",
    table.border.bottom.width = px(3),
    column_labels.border.bottom.color = "black",
    column_labels.border.bottom.width = px(2),
    heading.border.bottom.color = "#3778bf",
    heading.border.bottom.width = px(3)
  ) 
```

```{r}
# National totals of IMLS Outlets
imls_outlet_erate %>%
  select(funding_year, chosen_category_of_service, cat2_discount_by_ros, cat1_discount_by_ros_estimated) %>%
  filter(!is.na(funding_year)) %>% 
  group_by(funding_year) %>%
  summarise(Category1_Commitment = sum(cat1_discount_by_ros_estimated, na.rm=T),
            Category2_Commitment = sum(cat2_discount_by_ros, na.rm = T),
            National_Total = Category1_Commitment + Category2_Commitment)
```

```{r}
# National totals of IMLS AEs
imls_ae_erate %>%
  select(funding_year, chosen_category_of_service, cat2_discount_by_ros, cat1_discount_by_ros_estimated) %>%
  filter(!is.na(funding_year)) %>% 
  group_by(funding_year) %>%
  summarise(Category1_Commitment = sum(cat1_discount_by_ros_estimated, na.rm=T),
            Category2_Commitment = sum(cat2_discount_by_ros, na.rm = T),
            National_Total = Category1_Commitment + Category2_Commitment)
```

```{r}
# List all the entities
erate_libs %>%
  mutate(ros_entity_name = str_to_upper(ros_entity_name),
         ros_physical_city = str_to_upper(ros_physical_city)) %>%
  tidyr::unite(col = "lib_city_state", 
               c(ros_entity_name, ros_physical_city, ros_physical_state), 
               sep = ", ", 
               remove = FALSE) %>%
  distinct(lib_city_state) %>%
  arrange(lib_city_state)
```

```{r}
# Create a table of sum discounts that can be filtered by congressional district
erate_libs %>% 
  select(ros_physical_state, ros_congressional_district, funding_year, chosen_category_of_service, cat2_discount_by_ros, cat1_discount_by_ros_estimated) %>%
  mutate(ros_congressional_district = factor(ros_congressional_district, levels = stringr::str_sort(unique(ros_congressional_district), numeric = TRUE))) %>% 
  group_by(funding_year, ros_congressional_district, ros_physical_state) %>%
  summarise(sum_cat1_discount = sum(cat1_discount_by_ros_estimated, na.rm=T),
            sum_cat2_allocation = sum(cat2_discount_by_ros, na.rm = T),
            cong_dist_total = sum_cat1_discount + sum_cat2_allocation) %>%
  select(ros_congressional_district, ros_physical_state, funding_year, cong_dist_total) %>% 
  pivot_wider(names_from = funding_year, values_from = cong_dist_total) %>% 
  filter(ros_physical_state == "VI")
```

```{r}
erate_libs %>% 
  select(ros_entity_number, organization_entity_type_name, funding_year, chosen_category_of_service, cat2_discount_by_ros, cat1_discount_by_ros_estimated) %>%
  group_by(funding_year, organization_entity_type_name) %>%
  summarise(sum_cat1_discount = sum(cat1_discount_by_ros_estimated, na.rm=T),
            sum_cat2_allocation = sum(cat2_discount_by_ros, na.rm = T),
            org_type_total = sum_cat1_discount + sum_cat2_allocation,
            total_entities = n_distinct(ros_entity_number)) %>% 
  select(funding_year, organization_entity_type_name, org_type_total, total_entities) %>% 
  pivot_wider(names_from = organization_entity_type_name, values_from = org_type_total) #%>% 
  #write.csv("~/Google Drive File Stream/Shared drives/ALA-ERate/Data/Totals_by_Org_Type_by_Year.csv")
```

```{r}
erate_libs %>% 
  select(ros_entity_number, organization_entity_type_name, funding_year, chosen_category_of_service, cat2_discount_by_ros, cat1_discount_by_ros_estimated) %>%
  group_by(funding_year, organization_entity_type_name) %>%
  summarise(sum_cat1_discount = sum(cat1_discount_by_ros_estimated, na.rm=T),
            sum_cat2_allocation = sum(cat2_discount_by_ros, na.rm = T),
            org_type_total = sum_cat1_discount + sum_cat2_allocation,
            total_entities = n_distinct(ros_entity_number),
            per_entity_commitment = org_type_total/total_entities) %>% 
  select(funding_year, organization_entity_type_name, org_type_total, total_entities, per_entity_commitment) 
  #pivot_wider(names_from = organization_entity_type_name, values_from = c('org_type_total', 'total_entities', 'per_entity_commitment')) #%>% 
  #write.csv("~/Google Drive File Stream/Shared drives/ALA-ERate/Data/Totals_by_Org_Type_by_Year.csv")
```

```{r}
# Number of unique participants by entity type (library, lib system, or con) per year from 2016-2020
erate_libs %>% 
  select(ros_entity_number, organization_entity_type_name, funding_year) %>%
  group_by(funding_year, organization_entity_type_name) %>%
  summarise(total_entities = n_distinct(ros_entity_number)) %>% 
  pivot_wider(names_from = organization_entity_type_name, values_from = total_entities)
```

```{r}
# Number of unique library entities per year from 2016-2020
erate_libs %>% 
  select(ros_entity_number, funding_year) %>%
  group_by(funding_year) %>%
  summarise(total_entities = n_distinct(ros_entity_number)) 
```

```{r}
# Why do the total unique RENs differ when we group by org entity type or not?

```


```{r}
erate_libs %>%
  select(
    ros_entity_number,
    organization_entity_type_name,
    funding_year,
    chosen_category_of_service,
    cat2_discount_by_ros,
    cat1_discount_by_ros_estimated
  ) %>%
  group_by(funding_year, organization_entity_type_name) %>%
  summarise(
    sum_cat1_discount = sum(cat1_discount_by_ros_estimated, na.rm = T),
    sum_cat2_allocation = sum(cat2_discount_by_ros, na.rm = T),
    org_type_total = sum_cat1_discount + sum_cat2_allocation,
    total_entities = n_distinct(ros_entity_number),
    per_entity_commitment = org_type_total / total_entities
  ) %>%
  select(
    funding_year,
    organization_entity_type_name,
    org_type_total,
    total_entities,
    per_entity_commitment
  ) %>%
  ggplot(aes(x = funding_year, y = per_entity_commitment)) +
  geom_bar(
    aes(fill = reorder(organization_entity_type_name, -per_entity_commitment)),
    color = "black",
    size = 0.2,
    stat = "identity",
    position = "dodge"
  ) +
  theme_linedraw() +
  #scale_fill_jcolors(palette = "pal5") +
  scale_fill_manual(values = c("#3778bf", "#feb209", "#b0054b")) +
  scale_y_continuous(labels = scales::dollar_format()) +
  labs(title = "Per Recipient Funding Commitment by Organization Type",
       #subtitle = "IMLS City Size Designation & E-Rate Designation",
       fill = "Organization Type",
       x = "Funding Year",
       y = "Commitment Per Recipient")

#ggsave("~/Google Drive File Stream/Shared drives/ALA-ERate/Plots/All_Library_Recipient_Funding_by_Org_Type.png")
```

```{r}
erate_libs %>%
  select(
    ros_entity_number,
    organization_entity_type_name,
    funding_year,
    chosen_category_of_service,
    cat2_discount_by_ros,
    cat1_discount_by_ros_estimated
  ) %>%
  group_by(funding_year, organization_entity_type_name) %>%
  summarise(
    sum_cat1_discount = sum(cat1_discount_by_ros_estimated, na.rm = T),
    sum_cat2_allocation = sum(cat2_discount_by_ros, na.rm = T),
    org_type_total = sum_cat1_discount + sum_cat2_allocation,
    total_entities = n_distinct(ros_entity_number),
    per_entity_commitment = org_type_total / total_entities
  ) %>%
  select(
    funding_year,
    organization_entity_type_name,
    org_type_total,
    total_entities,
    per_entity_commitment
  ) %>%
  ggplot(aes(x = funding_year, y = per_entity_commitment)) +
  geom_bar(
    aes(fill = reorder(organization_entity_type_name, -per_entity_commitment)),
    color = "black",
    size = 0.4,
    stat = "identity",
    position = "dodge"
  ) +
  theme_linedraw() +
  scale_fill_manual(values = c("#484848", "#A9A9A9", "#DCDCDC")) +
  scale_y_continuous(labels = scales::dollar_format()) +
  labs(title = "Average Per Recipient Funding Commitment by Application Type",
       #subtitle = "IMLS City Size Designation & E-Rate Designation",
       fill = "Application Type",
       x = "Funding Year",
       y = "Commitment Per Recipient")

ggsave("~/Google Drive File Stream/Shared drives/ALA-ERate/Plots/Recipient_Funding_by_Org_Type.png")
```

```{r}
erate_libs %>%
  select(
    ros_entity_number,
    ros_urban_rural_status,
    funding_year,
    cat2_discount_by_ros,
    cat1_discount_by_ros_estimated
  ) %>%
  group_by(funding_year, ros_urban_rural_status) %>%
  summarise(
    sum_cat1_discount = sum(cat1_discount_by_ros_estimated, na.rm = T),
    sum_cat2_allocation = sum(cat2_discount_by_ros, na.rm = T),
    geog_type_total = round(sum_cat1_discount + sum_cat2_allocation, 2),
    total_entities = n_distinct(ros_entity_number),
    per_entity_commitment = round(geog_type_total / total_entities, 2)
  ) %>%
  select(
    funding_year,
    ros_urban_rural_status,
    geog_type_total,
    total_entities,
    per_entity_commitment
  )
```


```{r}
erate_libs %>%
  filter(!is.na(ros_urban_rural_status)) %>% 
  select(
    ros_entity_number,
    ros_urban_rural_status,
    funding_year,
    cat2_discount_by_ros,
    cat1_discount_by_ros_estimated
  ) %>%
  group_by(funding_year, ros_urban_rural_status) %>%
  summarise(
    sum_cat1_discount = sum(cat1_discount_by_ros_estimated, na.rm = T),
    sum_cat2_allocation = sum(cat2_discount_by_ros, na.rm = T),
    geog_type_total = sum_cat1_discount + sum_cat2_allocation,
    total_entities = n_distinct(ros_entity_number),
    per_entity_commitment = geog_type_total / total_entities
  ) %>%
  select(
    funding_year,
    ros_urban_rural_status,
    geog_type_total,
    total_entities,
    per_entity_commitment
  ) %>%
  ggplot(aes(x = funding_year, y = per_entity_commitment)) +
  geom_bar(
    aes(fill = reorder(ros_urban_rural_status, -per_entity_commitment)),
    color = "black",
    size = 0.4,
    stat = "identity",
    position = "dodge"
  ) +
  theme_linedraw() +
  scale_fill_manual(values = c("#484848", "#A9A9A9", "#DCDCDC")) +
  scale_y_continuous(labels = scales::dollar_format()) +
  labs(title = "Average Per Recipient Funding Commitment by Location Type",
       #subtitle = "IMLS City Size Designation & E-Rate Designation",
       fill = "Location Type",
       x = "Funding Year",
       y = "Commitment Per Recipient")

ggsave("~/Google Drive File Stream/Shared drives/ALA-ERate/Plots/Recipient_Funding_by_Geog.png")
```

```{r}
# Same as above but just for imls_outlets
imls_outlet_erate %>%
  select(
    ros_entity_number,
    organization_entity_type_name,
    funding_year,
    chosen_category_of_service,
    cat2_discount_by_ros,
    cat1_discount_by_ros_estimated
  ) %>%
  group_by(funding_year, organization_entity_type_name) %>%
  summarise(
    sum_cat1_discount = sum(cat1_discount_by_ros_estimated, na.rm = T),
    sum_cat2_allocation = sum(cat2_discount_by_ros, na.rm = T),
    org_type_total = sum_cat1_discount + sum_cat2_allocation,
    total_entities = n_distinct(ros_entity_number),
    per_entity_commitment = org_type_total / total_entities
  ) %>%
  select(
    funding_year,
    organization_entity_type_name,
    org_type_total,
    total_entities,
    per_entity_commitment
  ) %>%
  ggplot(aes(x = funding_year, y = per_entity_commitment)) +
  geom_bar(
    aes(fill = reorder(organization_entity_type_name, -per_entity_commitment)),
    color = "black",
    size = 0.2,
    stat = "identity",
    position = "dodge"
  ) +
  theme_linedraw() +
  scale_fill_manual(values = c("#3778bf", "#feb209", "#b0054b"), na.translate=FALSE) +
  scale_y_continuous(labels = scales::dollar_format()) +
  labs(title = "Per Recipient Funding Commitment by Organization Type",
       subtitle = "for IMLS Outlets that Received ERate Commitments",
       fill = "Organization Type",
       x = "Funding Year",
       y = "Commitment Per Recipient")

ggsave("~/Google Drive File Stream/Shared drives/ALA-ERate/Plots/IMLS_Outlet_Recipient_Funding_by_Org_Type.png")
```

```{r}
# How many entities are rural vs urban
erate_imls %>%
  group_by(ros_entity_number) %>%
  slice(1) %>%
  group_by(ros_urban_rural_status) %>%
  count()
```

```{r}
# How much is committed based on rural vs urban
erate_imls %>%
  group_by(funding_year, ros_urban_rural_status) %>%
  summarise(Cat1Total_Commitment = round(sum(cat1_discount_by_ros_estimated, na.rm=T), 1),
            Cat2Total_Commitment = round(sum(cat2_discount_by_ros, na.rm=T), 1),
            Total_Commitment = Cat1Total_Commitment + Cat2Total_Commitment) %>% 
  select(-Cat1Total_Commitment, -Cat2Total_Commitment) %>% 
  pivot_wider(names_from = ros_urban_rural_status, values_from = Total_Commitment) 
```

```{r}
# Visualize how the IMLS size designation compares to the E-Rate size designation
erate_imls %>%
  group_by(ros_entity_number) %>%
  arrange(funding_year) %>%
  slice(1) %>%
  group_by(LOCALE_ADD_DESCR, ros_urban_rural_status) %>%
  count(name = "CountEntities") %>%
  ggplot(aes(x = CountEntities, y = LOCALE_ADD_DESCR, fill = ros_urban_rural_status)) +
  geom_col(color = "black", size = 0.2) +
  theme_linedraw() +
  scale_fill_manual(values = c("#3778bf", "#feb209", "#b0054b"), na.value = "#d7d7ef") +
  labs(
    title = "Count of Entities",
    subtitle = "IMLS City Size Designation & E-Rate Designation",
    fill = "E-Rate Designation",
    caption = "NA = No Data Available"
  ) +
  xlab("Number of Entities") +
  ylab("IMLS Designation")

#ggsave("~/Google Drive File Stream/Shared drives/ALA-ERate/Plots/Entities_by_LOCALE_and_ROS_SIZE.png")
```

```{r}
# Cat 1 and Cat 2 Funding by IMLS Size for 2019
erate_imls %>%
  select(LOCALE_ADD_DESCR, funding_year, chosen_category_of_service, cat2_discount_by_ros, cat1_discount_by_ros_estimated) %>%
  filter(funding_year == "2019") %>%
  group_by(LOCALE_ADD_DESCR) %>%
  summarise(Category1_Commitment = round(sum(cat1_discount_by_ros_estimated, na.rm=T), 2),
            Category2_Commitment = round(sum(cat2_discount_by_ros, na.rm = T), 2),
            Total = round(Category1_Commitment + Category2_Commitment, 2)) %>%
  mutate(Pct_of_Lib_Funding = round((Total / sum(Total))*100, 2)) %>%
  arrange(desc(Pct_of_Lib_Funding))
```

```{r}
# Cat 1 and Cat 2 Funding by E-Rate Size for 2019
erate_imls %>% 
  select(ros_urban_rural_status, funding_year, chosen_category_of_service, cat2_discount_by_ros, cat1_discount_by_ros_estimated) %>%
  filter(funding_year == "2019") %>%
  group_by(ros_urban_rural_status) %>%
  summarise(Category1_Commitment = round(sum(cat1_discount_by_ros_estimated, na.rm = T), 2),
            Category2_Commitment = round(sum(cat2_discount_by_ros, na.rm = T), 2),
            Total = round(Category1_Commitment + Category2_Commitment, 2)) %>%
  mutate(Pct_of_Lib_Funding = round((Total / sum(Total))*100, 2)) %>%
  arrange(desc(Pct_of_Lib_Funding))
```

```{r}
# Total commitment by year by entity - currently filtered to IA
erate_libs %>%
  select(ros_entity_number, ros_entity_name, ros_physical_state, funding_year, 
         chosen_category_of_service, cat2_discount_by_ros, cat1_discount_by_ros_estimated) %>%
  filter(ros_physical_state == "IA") %>%
  group_by(ros_entity_number, ros_entity_name, funding_year) %>%
  summarise(Category1_Commitment = round(sum(cat1_discount_by_ros_estimated, na.rm = T), 2),
            Category2_Commitment = round(sum(cat2_discount_by_ros, na.rm = T), 2),
            Entity_Total = round(Category1_Commitment + Category2_Commitment, 2)) %>%
  select(ros_entity_number, ros_entity_name, funding_year, 
         Entity_Total) %>%
  arrange(funding_year) %>% 
  pivot_wider(names_from = funding_year, values_from = Entity_Total)
```

# How many of the IMLS outlets are participating in ERate?

```{r}
imls_outlet_erate %>%
  filter(STATSTRU != 3) %>% # filter out the closed libraries
  group_by(FSCSKEY, FSCS_SEQ) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(STABR) %>%
  summarise(Total_Count = n_distinct(FSCSKEY, FSCS_SEQ),
            ERate_Count = n_distinct(ros_entity_number, na.rm = T),
            NA_Count = sum(is.na(ros_entity_number)),
            Portion_Participating = ERate_Count / Total_Count) #%>%
  #write.csv("~/Google Drive File Stream/Shared drives/ALA-ERate/Data/2018IMLS_Outlets_Participating_in_ERate_by_State.csv")
```

```{r}
imls_ae_erate %>%
  filter(STATSTRU != 3) %>% # filter out the closed libraries
  group_by(FSCSKEY, FSCS_SEQ) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(STABR) %>%
  summarise(Total_Count = n_distinct(FSCSKEY, FSCS_SEQ),
            ERate_Count = n_distinct(ros_entity_number, na.rm = T),
            NA_Count = sum(is.na(ros_entity_number)),
            Portion_Participating = ERate_Count / Total_Count) #%>%
  #write.csv("~/Google Drive File Stream/Shared drives/ALA-ERate/Data/2018IMLS_AEs_Participating_in_ERate_by_State.csv")
```

# Disbursement Analysis

```{r}
# erate_libs %>%
#   left_join(disbursement %>% 
#               filter(form_version == "Current") %>%
#               select(application_number, funding_year, state, ben, organization_name, funding_request_number,
#                      form_471_service_type_name, form_version, total_authorized_disbursement, funding_commitment_request),
#             by = c("application_number", "organization_name", "funding_year", "funding_request_number", "form_471_service_type_name")
#             ) %>% 
#   distinct(application_number, funding_year, organization_name, form_471_service_type_name, .keep_all = T) %>% 
#   select(-ros_entity_number:-ros_number_of_nslp_students, -count_ros, -cat1_discount_by_ros_estimated, -cat2_discount_by_ros) %>% 
#   mutate(pct_of_committed_received = total_authorized_disbursement/post_discount_extended_eligible_line_item_costs) %>% 
#   filter(org_city == "OLYMPIA")
  #write.csv("~/Google Drive File Stream/Shared drives/ALA-ERate/Data/Lorenzo.csv")
```

Lorenzo's methodology:

To confirm the tables that you are using, I am assuming that the commitment table is the E-rate Recipient Details And Commitments dataset https://opendata.usac.org/E-rate/E-rate-Recipient-Details-And-Commitments/avi8-svp9 and the disbursement table is the E-rate Request for Discount on Services: FRN Status (FCC Form 471 and Related Information) dataset https://opendata.usac.org/E-rate/E-rate-Request-for-Discount-on-Services-FRN-Status/qdmp-ygft; if not, please let me know which ones you used.
 
Ensure correct level of aggregation calculations:
1. The disbursement table is at the FRN level while the commitment table is at the Recipient of service (ROS) FRN Line Item level. This means that in the commitment table an FRN would be replicated for each Line item in an FRN, and each line item would be replicated for each ROS in each line item. With this in mind, before you can join the tables, you have to ensure that your calculations using the right level of aggregation on the commitment table (roll up to the FRN total level):
  a. Sum the field post_discount_extended_eligible_line_item_costs for each FRN line item to get the total for the FRN (ensure you dedup by not including duplicate lines because of multiple ROSs for a given FRN line item)
  b. You can then use that sum to calculate
    i. pct_of_committed_received = total_authorized_disbursement / sum_ post_discount_extended_eligible_line_item_costs
 
Ensure you apply correct form and FRN status filters:
2. On the commitment table:
  a. form_471_status_name = Committed
  b. form_471_frn_status_name = funded (disbursements would only be authorized on funded FRNs)
3. On the disbursement table:
  a. form_version = Current (this means the form was processed and committed)
 
Joining fields: (later he clarified: the only joining field is Funding Request Number/FRN (funding year is not necessary as FRNs are unique))
1. funding_year
2. funding_request_number
 
Way to check correct joining:
1. You can verify if you are joining correctly by ensuring that for each FRN:
  a. [commitment table]. sum_ post_discount_extended_eligible_line_item_costs = [disbursement table]. funding_commitment_request
 
```{r}
# # this code is now implemented in the ec2 script that runs daily as disbursement_merged dataset
# disbursement_merged <- disbursement_merge %>%
#   distinct(funding_request_number, form_471_line_item_number, .keep_all = T) %>%
#   mutate(post_discount_extended_eligible_line_item_costs =
#            as.numeric(post_discount_extended_eligible_line_item_costs),
#          funding_request_number = as.numeric(funding_request_number)) %>%
#   group_by(funding_request_number) %>%
#   summarise(post_discount_extended_eligible_line_item_costs_sum =
#            sum(post_discount_extended_eligible_line_item_costs)) %>%
#   left_join(disbursement %>%
#               select(funding_request_number, funding_commitment_request, total_authorized_disbursement),
#             by = "funding_request_number"
#             ) %>%
#   mutate(pct_of_committed_received =
#            total_authorized_disbursement / post_discount_extended_eligible_line_item_costs_sum)
```

```{r}
# Use Lorenzo's test to see if the calculations and joins worked correctly
# You can verify if you are joining correctly by ensuring that for each FRN:
# a. [commitment table]. sum_ post_discount_extended_eligible_line_item_costs = [disbursement table].
# funding_commitment_request
isTRUE(disbursement_merged$post_discount_extended_eligible_line_item_costs_sum == disbursement_merged$funding_commitment_request)
```

```{r}
# Check if it's a rounding issue
disbursement_merged %>%
  #mutate(post_discount_extended_eligible_line_item_costs_sum =
           #base::ceiling(post_discount_extended_eligible_line_item_costs_sum),
         #funding_commitment_request = base::ceiling(funding_commitment_request)) %>% 
  filter(post_discount_extended_eligible_line_item_costs_sum - funding_commitment_request > 1 | post_discount_extended_eligible_line_item_costs_sum - funding_commitment_request < -1) 
  #write.csv("~/Google Drive File Stream/Shared drives/ALA-ERate/Data/Lorenzo_Sum_and_FCR_mismatch.csv")
```

```{r}
# Take a closer look at a FRN that is not matching as above chunk displays
read.csv("https://opendata.usac.org/resource/avi8-svp9.csv?funding_request_number=1799001258")
```

```{r}
# Check to see if any orgs were disbursed greater than 100% of committed funds
disbursement_merged %>%
  filter(pct_of_committed_received > 1.01) 
  #write.csv("~/Google Drive File Stream/Shared drives/ALA-ERate/Data/Lorenzo_Greater_than_1_Disbursement.csv")
```

From Lorenzo: This is to follow up on bullet point 1.a below: for those instances where there is a mismatch of more than $1, I confirmed that the correct FRN discounted total is in the funding_commitment_request field (from the FRN Status dataset). The calculation in that field is using the current discount rate.
 
For those 105 FRNs, you should use that field for your calculations as opposed to the sums from post_discount_extended_eligible_line_item_costs (from the E-rate Recipient Details And Commitments dataset), which is using a discount rate version that later changed during review. I submitted a ticket to have those records corrected.

```{r}
# Take out the FRNs that don't calculate correctly
# between function https://suzan.rbind.io/2018/02/dplyr-tutorial-3/#filtering-rows-based-on-a-numeric-variable
disbursement_merged %>% 
  # Keep NA's (this needs to be explicit in the filter call)
  filter(is.na(post_discount_extended_eligible_line_item_costs_sum - funding_commitment_request) |
           between(post_discount_extended_eligible_line_item_costs_sum - funding_commitment_request, -1, 1),
         is.na(pct_of_committed_received) | pct_of_committed_received < 1.01)
```

```{r}
special_frn_df <- setdiff(disbursement_merged %>% distinct(funding_request_number), 
               disbursement_merged %>% 
                 filter(is.na(post_discount_extended_eligible_line_item_costs_sum - funding_commitment_request) |
                          between(post_discount_extended_eligible_line_item_costs_sum - funding_commitment_request, -1, 1),
                        is.na(pct_of_committed_received) | pct_of_committed_received < 1.01) %>% 
          distinct(funding_request_number)
        )
```

```{r}
erate_libs %>% 
  filter(ros_physical_state != "PR",
         !is.na(form_471_frn_fiber_type_name))
```

```{r}
erate_libs %>% 
  filter(ros_physical_state != "PR",
         !is.na(form_471_frn_fiber_type_name)) %>% 
  summarise(n_distinct(ros_entity_number))
```

```{r}
erate_libs %>% 
  filter(ros_physical_state != "PR",
         form_471_function_name == "Fiber") 
```

```{r}
erate_libs %>% 
  filter(ros_physical_state != "PR",
         chosen_category_of_service == "Category1",
         form_471_function_name == "Fiber") %>% 
  summarise(n_distinct(ros_entity_number))
```
## C2 Budget
https://www.usac.org/e-rate/applicant-process/applying-for-discounts/category-two-budget/

```{r}
read.csv("https://opendata.usac.org/resource/6brt-5pbv.csv?$select=distinct%20applicant_type")
```


```{r}
# c2budget <- read.socrata(
#   "https://opendata.usac.org/resource/6brt-5pbv.json?$where=starts_with(applicant_type,%20'Library')",
#   app_token = Sys.getenv("USAC_Socrata")
# )
```


```{r}
## This code is incorrect because it does not add in the 20% allocation for 2020
# erate_imls <- erate_imls %>%
#   mutate(
#     c2_2015_to_2020_budget_rough = ifelse(
#       ros_entity_type != "non-instructional facility (nif)",
#       case_when(
#         !is.na(ros_square_footage) &
#           !is.na(LOCALE_ADD_DESCR) &
#           LOCALE_ADD_DESCR %in% c("City Large", "City Midsize", "Suburban Large") ~ ifelse(
#             ros_square_footage * 6.52 > 11998.43,
#             ros_square_footage * 6.52,
#             11998.43
#           ),!is.na(ros_square_footage) &
#           !is.na(LOCALE_ADD_DESCR) &
#           !LOCALE_ADD_DESCR %in% c("City Large", "City Midsize", "Suburban Large") ~ ifelse(
#             ros_square_footage * 3.00 > 11998.43,
#             ros_square_footage * 3.00,
#             11998.43
#           ),!is.na(ros_square_footage) &
#           is.na(LOCALE_ADD_DESCR) &
#           ros_urban_rural_status == "Urban" ~ ifelse(
#             ros_square_footage * 6.52 > 11998.43,
#             ros_square_footage * 6.52,
#             11998.43
#           ),!is.na(ros_square_footage) &
#           is.na(LOCALE_ADD_DESCR) &
#           ros_urban_rural_status == "Rural" ~ ifelse(
#             ros_square_footage * 3.00 > 11998.43,
#             ros_square_footage * 3.00,
#             11998.43
#           ),
#         is.na(ros_square_footage) &
#           !is.na(LOCALE_ADD_DESCR) &
#           LOCALE_ADD_DESCR %in% c("City Large", "City Midsize", "Suburban Large") ~ ifelse(SQ_FEET *
#                                                                                              6.52 > 11998.43, SQ_FEET * 6.52, 11998.43),
#         is.na(ros_square_footage) &
#           !is.na(LOCALE_ADD_DESCR) &
#           !LOCALE_ADD_DESCR %in% c("City Large", "City Midsize", "Suburban Large") ~ ifelse(SQ_FEET *
#                                                                                               3.00 > 11998.43, SQ_FEET * 3.00, 11998.43),
#         is.na(ros_square_footage) &
#           is.na(LOCALE_ADD_DESCR) &
#           ros_urban_rural_status == "Urban" ~ ifelse(SQ_FEET * 6.52 > 11998.43, SQ_FEET *
#                                                        6.52, 11998.43),
#         is.na(ros_square_footage) &
#           is.na(LOCALE_ADD_DESCR) &
#           ros_urban_rural_status == "Rural" ~ ifelse(SQ_FEET * 3.00 > 11998.43, SQ_FEET *
#                                                        3.00, 11998.43)
#       ),
#       NA
#     )
#   )
```

```{r}
# Calculate the pre-discount Category 2 funds requested by IMLS libraries
erate_imls %>% 
  filter(!is.na(FSCSKEY) & !is.na(FSCS_SEQ),
         chosen_category_of_service == "Category2",
         organization_entity_type_name %in% c("library system", "library")) %>% 
  add_count(funding_year, billed_entity_number, form_471_line_item_number, name = "entity_count") %>%
    # Add estimated amount received by individual recipient by calculating
  mutate(c2_prediscount_amt = as.numeric(pre_discount_extended_eligible_line_item_costs) /
             entity_count) %>% 
  summarise(sum(c2_prediscount_amt, na.rm = T))
```

```{r}
# Calculate the c2 budget in 2020 dollars for all IMLS entities
ae_and_outlets %>% 
  group_by(FSCSKEY, FSCS_SEQ) %>% 
  slice(which.max(SQ_FEET), with_ties = FALSE) %>% 
  ungroup() %>% 
  mutate(c2_budget = case_when(
    !is.na(SQ_FEET) & !is.na(LOCALE) & LOCALE %in% c(11,12,21) ~ ifelse(SQ_FEET*6.52 > 11998.43, SQ_FEET*6.52, 11998.43),
    !is.na(SQ_FEET) & !is.na(LOCALE) & !LOCALE %in% c(11,12,21) ~ ifelse(SQ_FEET*3.00 > 11998.43, SQ_FEET*3.00, 11998.43),
    !is.na(SQ_FEET) & is.na(LOCALE) & LOCALE_ADD %in% c(11,12,21) ~ ifelse(SQ_FEET*6.52 > 11998.43, SQ_FEET*6.52, 11998.43),
    !is.na(SQ_FEET) & is.na(LOCALE) & !LOCALE_ADD %in% c(11,12,21) ~ ifelse(SQ_FEET*3.00 > 11998.43, SQ_FEET*3.00, 11998.43)
    )) %>% 
  summarise(sum(c2_budget, na.rm = T))
```

```{r}
# Calculate the pre-discount Category 2 funds requested by IMLS libraries
erate_imls %>% 
  filter(!is.na(FSCSKEY) & !is.na(FSCS_SEQ),
         chosen_category_of_service == "Category2",
         organization_entity_type_name %in% c("library system", "library")) %>% 
  group_by(ros_entity_number) %>% 
  slice_head(n = 1) %>% 
  ungroup() %>% 
  summarise(TotalBudget = sum(c2_2015_to_2020_budget_rough, na.rm = T))
```

Question of the day: What percentage of recipients of Category 1 commitments used that funding for fiber? % per funding year, also by state

form_471_function_name

```{r}
erate_libs %>% 
  filter(chosen_category_of_service == "Category1") %>% 
  group_by(funding_year) %>% 
  summarise(n_distinct(ros_entity_number))
```

```{r}
erate_libs %>%
  filter(chosen_category_of_service == "Category1") %>%
  count(ros_entity_number, funding_year, form_471_function_name, form_471_frn_fiber_type_name) %>%
  filter(form_471_function_name == "Fiber") %>%
  group_by(funding_year) %>%
  summarise(num_committed_for_fiber = n_distinct(ros_entity_number)) %>%
  left_join(
    erate_libs %>%
      filter(chosen_category_of_service == "Category1") %>%
      group_by(funding_year) %>%
      summarise(num_committed_for_cat1 = n_distinct(ros_entity_number))
  ) %>% 
  mutate(portion_cat1_fiber = num_committed_for_fiber/num_committed_for_cat1) 
```


```{r}
erate_libs %>%
  filter(chosen_category_of_service == "Category1") %>%
  count(ros_entity_number, funding_year, form_471_function_name) %>%
  filter(form_471_function_name == "Fiber") %>%
  group_by(funding_year) %>%
  summarise(num_committed_for_fiber = n_distinct(ros_entity_number)) %>%
  left_join(
    erate_libs %>%
      filter(chosen_category_of_service == "Category1") %>%
      group_by(funding_year) %>%
      summarise(num_committed_for_cat1 = n_distinct(ros_entity_number))
  ) %>% 
  mutate(portion_cat1_fiber = num_committed_for_fiber/num_committed_for_cat1) %>% 
  ggplot(aes(x=funding_year, y=portion_cat1_fiber)) +
  geom_bar(color = "black", size = 0.2, fill="#3778bf", stat = "identity") +
  theme_linedraw() +
  geom_text(aes(label = scales::percent(portion_cat1_fiber)), vjust = 1.5, colour = "white") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Of the libraries that were committed Category 1 funding,\nwhat percentage were committed funds for fiber?") +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())
```

```{r}
erate_libs %>%
  filter(chosen_category_of_service == "Category1",
         form_471_function_name == "Fiber") %>%
  group_by(funding_year, form_471_function_name, form_471_frn_fiber_type_name, form_471_frn_fiber_sub_type_name) %>% 
  summarise(Category1_Commitment = if (all(is.na(cat1_discount_by_ros_estimated)))
      NA_real_
    else
      scales::dollar(sum(cat1_discount_by_ros_estimated, na.rm = T), largest_with_cents = 10^9))
```

The state report - can you run it so that differences of greater than 10% (+/-) per category when comparing year-to-year are pulled out into a summary? (ahhhhhhh - getting distracted by e-rate - must turn away) Thinking if we had an overview page that showed changes in 2016 per state for any of those gains - Table lists all the states that had a change (and which way +/-) by category, then a table for 2017, and so on

https://www.infoworld.com/article/3404276/how-to-calculate-month-over-month-changes-in-r.html

```{r}
erate_libs %>% 
  select(ros_physical_state, 
         funding_year,
         chosen_category_of_service,
         cat2_discount_by_ros,
         cat1_discount_by_ros_estimated) %>%
  group_by(ros_physical_state, funding_year) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(Category1_Commitment = if(all(is.na(cat1_discount_by_ros_estimated))) NA_real_ 
            else sum(cat1_discount_by_ros_estimated, na.rm = T),
            Category2_Commitment = if(all(is.na(cat2_discount_by_ros))) NA_real_ 
            else sum(cat2_discount_by_ros, na.rm = T)) %>% 
  arrange(ros_physical_state, funding_year) %>% 
  mutate(YoY_Cat1 = (Category1_Commitment - lag(Category1_Commitment)) / lag(Category1_Commitment),
         YoY_Cat2 = (Category2_Commitment - lag(Category2_Commitment)) / lag(Category2_Commitment)) 
```

```{r}
erate_libs %>% 
  select(ros_physical_state,
         ros_entity_number,
         funding_year,
         chosen_category_of_service,
         cat2_discount_by_ros,
         cat1_discount_by_ros_estimated) %>%
  group_by(ros_physical_state, funding_year) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(Category1_Commitment = if(all(is.na(cat1_discount_by_ros_estimated))) NA_real_ 
            else sum(cat1_discount_by_ros_estimated, na.rm = T),
            Category2_Commitment = if(all(is.na(cat2_discount_by_ros))) NA_real_ 
            else sum(cat2_discount_by_ros, na.rm = T),
            num_entities = n_distinct(ros_entity_number)) %>% 
  rowwise() %>% 
  mutate(National_Total = sum(c(Category1_Commitment, Category2_Commitment), na.rm = TRUE),
         per_entity_avg = National_Total/num_entities) %>% 
  ungroup() %>% 
  group_by(ros_physical_state) %>% 
  arrange(ros_physical_state, funding_year) %>%
  mutate(YoY_Cat1 = (Category1_Commitment - lag(Category1_Commitment)) / lag(Category1_Commitment),
         YoY_Cat2 = (Category2_Commitment - lag(Category2_Commitment)) / lag(Category2_Commitment),
         YoY_Num = (num_entities - lag(num_entities)) / lag(num_entities),
         YoY_EntAvg = (per_entity_avg - lag(per_entity_avg)) / lag(per_entity_avg)) %>% 
  select(ros_physical_state, funding_year, YoY_Cat1:YoY_EntAvg) %>% 
  pivot_longer(YoY_Cat1:YoY_EntAvg, names_to = "YoY_Comparison", values_to = "Change_from_Prev_Year") %>% 
  filter(funding_year == 2019,
         abs(Change_from_Prev_Year) > 0.50) %>% 
  mutate(Change_from_Prev_Year = scales::percent(Change_from_Prev_Year)) 
  
```


```{r}
erate_libs %>% 
  filter(ros_physical_state == "WA",
         funding_year < 2020) %>% 
  select(ros_physical_state,
         ros_entity_number,
         funding_year,
         chosen_category_of_service,
         cat2_discount_by_ros,
         cat1_discount_by_ros_estimated) %>%
  group_by(ros_physical_state, funding_year) %>%
  # Code within summarise is to ensure we keep NAs and that they don't become zeroes
  summarise(Category1_Commitment = if(all(is.na(cat1_discount_by_ros_estimated))) NA_real_ 
            else sum(cat1_discount_by_ros_estimated, na.rm = T),
            Category2_Commitment = if(all(is.na(cat2_discount_by_ros))) NA_real_ 
            else sum(cat2_discount_by_ros, na.rm = T),
            num_entities = n_distinct(ros_entity_number)) %>% 
  rowwise() %>% 
  mutate(National_Total = sum(c(Category1_Commitment, Category2_Commitment), na.rm = TRUE),
         per_entity_avg = National_Total/num_entities) %>% 
  ungroup() %>% 
  group_by(ros_physical_state) %>% 
  arrange(ros_physical_state, funding_year) %>%
  mutate(YoY_Cat1 = (Category1_Commitment - lag(Category1_Commitment)) / lag(Category1_Commitment),
         YoY_Cat2 = (Category2_Commitment - lag(Category2_Commitment)) / lag(Category2_Commitment),
         YoY_Num = (num_entities - lag(num_entities)) / lag(num_entities),
         YoY_EntAvg = (per_entity_avg - lag(per_entity_avg)) / lag(per_entity_avg)) %>% 
  select(ros_physical_state, funding_year, YoY_Cat1:YoY_EntAvg) %>% 
  pivot_longer(YoY_Cat1:YoY_EntAvg, names_to = "YoY_Comparison", values_to = "Change_from_Prev_Year") %>% 
  ggplot(aes(x = funding_year, y = Change_from_Prev_Year, color = YoY_Comparison)) +
  geom_point() +
  geom_line() +
  theme_linedraw() +
  geom_text(aes(label = scales::percent(Change_from_Prev_Year)), vjust = 1.5, colour = "black") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Year over year percentage change") +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())
  
```


